---
title: "schisto_inlabru_V1"
author: "Eberle Yarborough"
date: "`r Sys.Date()`"
output: pdf_document
---

This document is basic exploration of the data to be included in the *Schistosoma* spp. spatial Bayesian analysis, before running a full INLA model. Here, we will use `inlabru` to explore the schisto data and determine if there is any spatial dependency.

**Previous Model Runs**
  - 3.5.26:
    - Mesh resolution = 25km
    - Range prior = 37, 0.05
    - sigma_u prior = 2, 0.05
    - Random effects = SRF only
    - Env covariates: TWI, MSI, EVI, clay %, weighted distance to fire
      - Smoothers: weighted distance to fire - RW1
    - Models:
      - Bernoulli GAM: env + demographics
      - Bernoulli GAM: env + demographics + SRF
      - Bernoulli GLM: SRF + % clay
    - Lowest WAIC model = SRF + clay % only
  - 3.6.26:
    - Mesh resolution = 15km
    - Range prior = 25, 0.05
    - sigma_u prior = 2.5, 0.05
    - Random effects = SRF only
    - Env covariates = TWI, MSI, EVI, clay %, weighted distance to fire
      - Smoothers: weighted distance to fire - RW2
    - Models:
      - Bernoulli GAM: env + demographics
      - Bernoulli GAM: env + demographics + SRF
      - Bernoulli GLM: SRF + age (years)
    - Lowest WAIC model = SRF and age only
  - 3.10.26:
    - Mesh resolution = 15km
    - Range prior = 25, 0.95
    - sigma_u prior = 2, 0.05
    - Random effects = SRF only
    - Env covariates = TWI, MSI, EVI, clay %, weighted distance to fire
      - Smoothers: all env covariates, RW2
    - Models:
      - Bernoulli GAM: env + demographics
      - Bernoulli GAM: env + demographics + SRF
      - Bernoulli GLM: SRF + log1p(age (years))
  - 3.17.26:
    - Mesh resolution = 15km
    - Range prior = 25, 0.95
    - sigma_u prior = 2, 0.05
    - Random effects = SRF only
    - Env covariates = TWI, MSI, EVI, clay %, weighted distance to fire
      - Smoothers: TWI, MSI, weighted distance to fire - all RW2
    - Models:
      - Bernoulli GAM: env + demographics
      - Bernoulli GAM: env + demographics + SRF
      - Bernoulli GLM: SRF + log1p(age (years))

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
  include = TRUE,
	message = FALSE,
	warning = FALSE
)

#' We need the following packages:
library(lattice)
library(ggplot2)
library(sf)
library(gstat)
#' If you've updated R recently, you have to re-install INLA from their website
#' This is for the stable version:
#install.packages("INLA",repos=c(getOption("repos"),INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(INLA)
library(inlabru)
library(DHARMa)
library(dplyr)
library(glmmTMB)
library(rnaturalearth)
library(scales)
library(cowplot)
library(tidyr)
source("/Users/eberleyarborough/Documents/Coding Help/INLA tutorials/highland_stats_INLA_course/AllRCodeInlabru/HighstatLibV14.R") #' Highland Statistics support file

#' Set the working directory and import the data (should be the project directory).
setwd("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Schisto/Schisto_Analysis")
```

### Section 1: Importing the Data & Exploration -----

The question we are attempting to answer is "What are the drivers of *Schistosoma* spp. infection in African buffalo (*Syncerus caffer*) within Kruger National Park (KNP)?" We have:
 - Host demographics: age (factor with 3 levels OR continuous), sex (factor with 2 levels), average body condition score (discrete)
 - Environmental covariates: preceding wet season (Jan 1 - Apr 30, 2023) enhanced vegetation index (EVI) and moisture stress index (MSI); proportion of clay in the soil up to 30cm in depth; weighted distance to pre-capture fires (Apr 2022 - March 2023); and the topographical wetness index (TWI) - a measure of water accumulation and flow on the landscape
 - Locations: 2023 buffalo mass-capture (cross-sectional) locations across KNP

```{r importing-and-transforming-data}
#' Step 1: Import the disease data -----
#' This file contains the schisto results, as well as the buffalo demographics,
#' capture locations, and elevation at those capture locations:
schisto_base<- read.csv("schisto_with_cap_demo_and_elev.csv")

#' converting to a spatial object
schisto.sf <- st_as_sf(schisto_base,
                  coords = c("x.east", "y.south"),
                  crs    = 32736)

#' Setting the CRS, converting to KM (it is best to work with small numbers in inlabru)
crs.Target <- fm_crs("EPSG:32736")  # Target for KNP in UTM 36S
CRS.km     <- fm_crs_set_lengthunit(crs.Target, "km") # Converting to km instead of m

#' We now transform the spatial object to the kilometre-based coordinate 
#' system and extract the coordinates as numeric variables. These will 
#' be used later when building spatial models.
schisto2.sf <- st_transform(schisto.sf, CRS.km)
Coords <- st_coordinates(schisto2.sf)
schisto2.sf$Xkm <- Coords[, 1]
schisto2.sf$Ykm <- Coords[, 2]

#' Dropping unnessary columns from the schisto data frame
schisto2.sf <- schisto2.sf %>%
  select("animal.ID", "age.years", "age.class", "Sex", "BCS.Average", "Result", "Xkm", "Ykm", "geometry")

#' Step 2: Import the environmental covariates -----
#' Read in the .csv file with the environmental covariates. NOTE: This file
#' contains data for all capture locations, not just those that are schisto positive.
env_data <- read.csv("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/CSVs/cap_and_env_data.csv")
env_data <- env_data %>%
  select("animal.ID", "capture.date", "capture.day", "capture.group.by.day", "approx.herd.size", "weighted_dist_to_fire", "topo_wetness_index", "clay_percent_30cm", "jan_apr_23_evi", "jan_apr_23_msi")

#' Join to schisto data
schisto.complete.sf <- left_join(schisto2.sf, env_data, by = "animal.ID")

#' Step 3: Transform Data

#' It's good to convert the character "positive/negative" results to 
#' something binary and numeric (0/1, where 1 is positive)
schisto.complete.sf <- schisto.complete.sf %>%
  mutate(result01 = ifelse(Result == "positive", 1, 0))

#' And convert Sex, age class, and capture group by day to factors
schisto.complete.sf$fage.class <- factor(schisto.complete.sf$age.class)
schisto.complete.sf$fSex <- factor(schisto.complete.sf$Sex)
schisto.complete.sf$fcapgroup <- factor(schisto.complete.sf$capture.group.by.day)

#' We'll also log1p transform age in case we want to model with it later on. We
#' also use the `log1p()` function here instead of just `log()` because we would
#' end up zero-inflating our dataset by taking the ln of 1 ( = 0).
schisto.complete.sf <- schisto.complete.sf %>%
  mutate(log.age = log1p(age.years))

#' Step 3: Import the KNP boundary and match CRS -----
#' To provide geographic context, we load a map of Kruger that will be 
#' used to visualize the study area and the sampling locations.
kruger_shape <- st_read("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/SANParks KNP Shapefiles/KNP_polygon/KNP_polygon.shp")

#' Transform to UTM Zone 36S in km
kruger.km <- st_transform(kruger_shape, crs = CRS.km)

#' Finally, we plot the KNP boundary together with the spatial distribution of 
#' the capture locations to confirm that the data are correctly georeferenced. 
ggplot() +
  geom_sf(data = kruger.km, fill = "transparent") +
  geom_sf(data = schisto.complete.sf) +
  theme_minimal()

#' Confirm that the data and the map share the same CRS: 
st_crs(kruger.km) == st_crs(schisto.complete.sf)
```

First, we'll define all of our mesh components and construct the mesh we'll use for the spatial random field (SRF).

```{r building-the-mesh}
#' Step 1: Making the mesh -----

#' To get a good idea of mesh settings for the SRF, we first need to get a sense what the distances are between the 
#' sampling locations. 
Loc <- cbind(schisto.complete.sf$Xkm, schisto.complete.sf$Ykm)
D   <- dist(Loc)
par(mfrow = c(1,1), mar = c(5,5,2,2))
hist(D, 
     freq = TRUE,
     main = "", 
     xlab = "Distance between sites (km)",
     ylab = "Frequency")
#' Sites are anywhere from 0-350km apart, so anything small scale is probably < 50 km.

#' The main tools to control the mesh are:
#' - max.edge: The largest allowed triangle length. The lower the value 
#'             for max.edge the higher the resolution. This should be "small-scale"
#'             based off of your sampling locations (1/5 the resolution).
#' - cutoff:   Sites with a distance value smaller than the cutoff 
#'             are modeled by a single vertex in the mesh.
#'             Set the cutoff to 1/5 of the maxedge of the inner area.
#' - boundary: Useful for islands and fjords, but you can make it hug your sampling
#'             locations tighter by making it nonconvex.
#  - offset:   Determines how far the inner and outer boundaries extend.
#' - Inner area / outer areas:
#'             Is to avoid potential numerical estimation problems for 
#'             the points close to the boundary. General recommendation is: 1:5.
#'             This explains the:   max.edge = c(1, 5) * MaxEdge in the 
#'             code below.

#' Now we should define boundaries for the spatial mesh.
#' First, define the resolution of the mesh based off of our distances between sites.
#' This can be adjusted to get at finer-scale (local) spatial patterns. 
MeshResolution <- 15               #' in km.
MaxEdge  <- MeshResolution / 5    #' good rule of thumb is to have max edge set to 1/5 the resolution

#' #' fm_nonconvex_hull makes a non-convex area around the sampling
#' locations - the inner boundary. You can control how close the blue line is 
#' to the sites with the 'convex' argument. Larger numbers here mean the inner
#' hull is less tightly bound to the capture sites. 
inner.hull <- fm_nonconvex_hull(Loc, convex = -0.07, crs = st_crs(schisto.complete.sf))
#' Results
plot(inner.hull)
points(Loc)

#' Convert the inner, nonconvex hull to an sf object
inner.hull_sf <- st_as_sf(inner.hull)

#' Then we make an outer ring by buffering the inner hull by about one
#' times the expected range. That distance is big enough that, by the 
#' time the field reaches the outer edge, most correlation has died 
#' away—so boundary effects are small. We simplify the outer boundary
#' a little; we don’t need every little curve for meshing purposes.

outer_sf   <- st_buffer(inner.hull_sf, 
                        dist = MaxEdge * 5)  
outer_sf_simple <- st_simplify(outer_sf,
                               dTolerance = 0.05 * MaxEdge * 5,
                               preserveTopology = TRUE)

#' Plot inner and outer boundaries
ggplot() +
  geom_sf(data = outer_sf_simple, 
          col = "red",  fill = NA, alpha = 0.5) +
  geom_sf(data = inner.hull_sf,    col = "blue", fill = NA) +
  geom_sf(data = kruger.km, fill = NA, alpha = 0.5) +
  geom_sf(data = schisto.complete.sf, color = "black", size = 0.75) +
  theme_minimal() +
  labs(x = "Easting (km)", y = "Northing (km)")

#' Before we can use the inner and outer boundaries in fm_mesh_2d(), we 
#' need to convert the two sf polygons into their outlines (the lines 
#' that trace around the polygons). The mesher does not take filled 
#' shapes; it only works with these outlines, which it represents 
#' as boundary segments.
inner.hull_segm <- fm_as_segm(inner.hull_sf)
outer.hull_segm <- fm_as_segm(outer_sf_simple)

#' Make the mesh with the new boundary segments
mesh1 <- fm_mesh_2d_inla(
  loc      = Loc,
  boundary = list(inner.hull_segm, outer.hull_segm),
  max.edge = c(1, 5) * MaxEdge,  
  cutoff   = MaxEdge / 5)
fm_crs(mesh1) <- st_crs(schisto.complete.sf)  

#' Size of mesh
mesh1$n 
#' 7071 nodes for 557 observations - decent. We can to set the resolution to 
#' finer-scale if we'd like.

#' Ensure the mesh has the same CRS as KNP 
fm_crs(mesh1) <- st_crs(kruger.km)  

#' Plot the study area on the mesh
knp.mesh.plot <- ggplot() +
            theme_minimal() +
            labs(title = "Study area in UTM (km)") +
            geom_fm(data = mesh1) +
            geom_sf(data = schisto.complete.sf,
                    alpha = 0.5,
                    size = 0.5)  +
            geom_sf(data = kruger.km,
                    fill = "transparent", 
                    col = "red") 
knp.mesh.plot
```

Now we'll explore the data a bit more to look at any underlying relationships, nonlinear patterns, etc.

```{r data-exploration}
#' Step 1: Determine if there are any missing values
MyVar <- c("animal.ID", "age.years", "log.age", "age.class", "Sex", "BCS.Average", "result01",
           "capture.date", "capture.day", "capture.group.by.day", "approx.herd.size",
           "weighted_dist_to_fire", "topo_wetness_index", "clay_percent_30cm",
           "jan_apr_23_evi", "jan_apr_23_msi")
colSums(is.na(schisto.complete.sf[,MyVar]))
#' Looks like missing values in herd size, EVI, and MSI (11 in each).

#' We'll drop NA rows
#' Now we'll remove any NA columns
schisto.complete.sf <- schisto.complete.sf %>%
  drop_na()
#' Left with 557 observations

#' For some analyses, it is convenient to work with a standard data frame 
#' rather than an sf object. We therefore drop the geometry column and 
#' store the result separately.
schisto.complete.df <- st_drop_geometry(schisto.complete.sf)

#' It's also a good idea to standardize continuous covariates - this can only be done after NA is dropped.
#' The `MyStd()` function is a custom function in the Highland Statistics 
#' script.
schisto.complete.sf$weighted_dist_to_fire.std <- MyStd(schisto.complete.sf$weighted_dist_to_fire)
schisto.complete.sf$topo_wetness_index.std <- MyStd(schisto.complete.sf$topo_wetness_index)
schisto.complete.sf$clay_percent_30cm.std <- MyStd(schisto.complete.sf$clay_percent_30cm)
schisto.complete.sf$jan_apr_23_evi.std <- MyStd(schisto.complete.sf$jan_apr_23_evi)
schisto.complete.sf$jan_apr_23_msi.std <- MyStd(schisto.complete.sf$jan_apr_23_msi)

#' Step 2: Look the depth of the data - do we have enough observations with the
#' given covariates?

#' How many observations per date?
table(schisto.complete.sf$capture.date)
#' Anywhere from 5 to 29

#' How many groups per day?
table(schisto.complete.sf$capture.group.by.day)
#' as much as 5 per day - consider using as a random effect?

#' What's the range of herd sizes?
hist(schisto.complete.sf$approx.herd.size)
#' looks like 50 - 600 animals per herd - consider using as a random effect?

#' Are herd size and capture group per day related?
plot(x = schisto.complete.sf$capture.group.by.day,
     y = schisto.complete.sf$approx.herd.size,
     xlab = "Capture Group by Day",
     ylab = "Approx. Herd Size")
#' fairly even spread of herd sizes across capture groups - pick one to use
#' as a random effect, not both

#' How many per age class?
table(schisto.complete.sf$age.class)
#' 71 calves, 387 subadults, 96 adults, 3 old adults - would need to combine
#' adults and old adults if wanting to use this as a covariate or as an 
#' interaction term

#' What is the age range?
hist(schisto.complete.sf$age.years)
hist(schisto.complete.sf$log.age)
range(schisto.complete.sf$age.years)
#' as young as 0.5 to 20.27 years

#' What are the differences in sex?
table(schisto.complete.sf$Sex) #' enough observations between sexes
table(schisto.complete.sf$Sex, schisto.complete.sf$result01) #' no separation
#' there are positive and negative males and females - no category has 0 observations,
#' which means we can use a Bernoulli distribution of the response variable.
#' What about with age?
table(schisto.complete.sf$age.class, schisto.complete.sf$result01)
#' The only issue we have here is with "old adult," but we already said we would
#' likely combine with "adult" since there are only 3 observations of old adults,
#' anyway.

#' Step 3: Check for outliers - the `MyDotplot.ggp2()` function comes from the
#' Highland Statistics custom script.
MyDotplot.ggp2(schisto.complete.df, MyVar, Ncol = 3, TextSize = 10,
               PointSize = 0.5, DropLabels = FALSE)
#' No obvious outliers, but possible collinearity or non-linear patterns.

#' Step 4: Assess for collinearity
#' Make a pairplot - the `Mypairs()` function comes from the Highland Statistics
#' custom script.
MyVar <- c("Xkm", "Ykm", "log.age", "BCS.Average", "clay_percent_30cm",
           "jan_apr_23_evi", "jan_apr_23_msi", "topo_wetness_index",
           "weighted_dist_to_fire")
Mypairs(schisto.complete.df[,MyVar])
#' There might be some collinearity between Xkm and/or Ykm and percentage of clay
#' in the soil, as well as EVI and MSI, which makes sense, as these are spatial
#' covariates.

#' We only should be nervous if a Pearson correlation is larger than 0.8; however,
#' that doesn't mean we should ignore moderately high (~0.6) values, like
#' those for X/Ykm and clay percentage.

#' If we had temporal (longitudinal) data, here's where we would check to see how/if
#' covariates change over time by plotting every covariate vs. year.

#' Do covariates differ by clay percentage in the soil? We anticipate there 
#' may be a relationship, as the clay content influences water retention
#' on the landscape (MSI) and vegetation growth (EVI).
#' First, we need to put the data into "long format" for easier plotting.
GatherTheseCols <- c("jan_apr_23_msi", "jan_apr_23_evi", "topo_wetness_index",
                     "weighted_dist_to_fire", "log.age")
schist_long <- gather(data = schisto.complete.df,
                      key = "ID",
                      value = "AllX",
                      all_of(GatherTheseCols),
                      factor_key = TRUE)
dim(schist_long)

#' Check to see how much variation in EVI/MSI is explained by clay percentage
#' with SLR.
test_msi <- lm(jan_apr_23_msi ~ clay_percent_30cm, data = schisto.complete.sf)
summary(test_msi)
#' approximately 1.8%

test_evi <- lm(jan_apr_23_evi ~ clay_percent_30cm, data = schisto.complete.sf)
summary(test_evi)
#' Less than 1%

#' What about variation in MSI explained by EVI?
test_msi_evi <- lm(jan_apr_23_msi ~ jan_apr_23_evi, data = schisto.complete.sf)
summary(test_msi_evi)
#' About 25%, so not insignificant, but shouldn't throw too many issues in the model


#' Step 5: Continuous covariates vs. disease data
#' Plot each continuous covariate against the presence/absence data. Note:
#' the disease data MUST be in numeric 0/1 form to see if relationships
#' are non-linear with this plot.
ggplot(data = schist_long, aes(x=AllX, y = result01)) +
  geom_point(size = 1, alpha = 0.5) +
  labs(x = "Covariate", y = "Disease Status") +
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 5)) +
  geom_smooth(method = "gam",
              method.args = list(family = "binomial"),
              se = FALSE) +
  facet_wrap(.~ID, scale = 'free_x', ncol = 2)
#' Only non-linear pattern appears to be weighted distance to fire, so 
#' we might need a smoother for this covariate.

#' Step 6: Assess for dependency
#' Here are the spatial patterns of the capture locations (using base R)
xyplot(Ykm ~ Xkm,
       aspect = "iso",
       col = 1,
       pch = 16,
       data = schisto.complete.sf)

xyplot(Ykm ~ Xkm | fcapgroup,
       aspect = "iso",
       col = 1,
       pch = 16,
       data = schisto.complete.sf)
#' It looks like extra captures (like 1a 1b, 2a 2b) happened in the north,
#' and up to 5 captures per day happened in the southeast, near Lower Sabie.
#' Most capture days had only 2 captures per day. Just something to think about
#' if we wanted to include capture group by day in the analyses in any way. We 
#' would probably remove 1a/b and 2a/b since there's no real variation there, and
#' we would just call them "1" and "2."

#' Here's the same kind of plot as a ggplot
ggplot(data = kruger.km) + #' boundary
  geom_sf(fill = "transparent") + #' no fill
  geom_sf(data = schisto.complete.sf, #' capture locations
          size = 0.75, #' size of the points
          alpha = 0.5) + #' transparency of the points
  theme_minimal() +
  labs(title = "Study Area in UTM (km)")

#' And split up by capture group
ggplot(data = kruger.km) +
  geom_sf(fill = "transparent") +
  geom_sf(data = schisto.complete.sf,
          size = 0.75,
          alpha = 0.5) +
  theme_minimal() +
  labs(title = "Study Area in UTM (km)") +
  facet_wrap(~fcapgroup)
```

To conclude about our data exploration:
- We have spatial patterns in some of our covariates, which isn't surprising.
- We don't have any quasi-complete separation in the pos/neg disease data.
- Herd size or capture group could be used as random effects; however, if using
capture group by day, we should collapse 1a/b and 2a/b into 1 and 2.
- If we include weighted distance to fire, it should probably be a GAM.
- We can consider smoothers for the other environmental covariates.
- If we want to use age class, we should collapse "adult" and "old adult" since
there are only 3 "old adult" animals.
- It doesn't look like we have any dependency (other than potential spatial
dependency).

### Section 2: Model Components -----

#### Subsection 2.1: Formulating the Bernoulli GAM

There are two ways to end up with a Bernoulli GLM/GAM for these data.
1. If your question is: What is driving presence/absence of schisto,
    then you can go straight to a Bernoulli GLM/GAM.
2. The abundance data have a high % of zeros, and a Poisson GLM cannot
    cope with this high % of zeros. It is an option (among many) to 
    apply a hurdle model. In such a model, we analyse the
    absence/presence data with a Bernoulli GLM/GAM, the presence-only
    data with a truncated Poisson GLM/GAM and then glue them together. 
    You might do this analysis *after* looking at the one in option 1 -
    after you've figured out what is driving presence/absence, then you
    can look at abundance.

Here, we'll apply the following equation:

**Model 1:**
  result01_i ~ Bernoulli(Pi_i)
  E[result01] = Pi_i
  var[result01] = Pi_i * (1 - Pi)
  
With our covariates, this shakes out to:
  logit(Pi_i) = Intercept + age.years_i + BCS.Average_i + s(weighted_dist_to_fire.std) +
                topo_wetness_index.std_i + clay_percent_30cm.std_i +
                jan_apr_23_evi.std_i + jan_apr_23_msi.std_i 

Alternatively, with the log1p age and the smoother terms for all environmental 
covariates:
  logit(Pi_i) = Intercept + log.age_i + BCS.Average_i + f(weighted_dist_to_fire) +
                f(topo_wetness_index) + f(clay_percent_30cm) +
                f(jan_apr_23_evi.std) + f(jan_apr_23_msi) 

There's also the option of including herd size or capture group as a random
effect. But we can do that in later models.

We can write the logit link function also as:

           exp(Intercept + Covariates_i)  
  Pi_i = ---------------------------------------------------------------------------
         1 + exp(Intercept + Covariates_i)  


**Model 2:**
This is the same as model 1 (Bernoulli GAM) but with spatial dependency.

  logit_(Pi_i) = Intercept + Covariates + u_i, where u_i are the spatially
  correlated random effects
  
We can write this logit link function as: 

           exp(Intercept + Covariates_i + u_i)  
  Pi_i = ------------------------------------------
         1 + exp(Intercept + Covariates_i+ u_i)  


**Model 3:**
This model will assess whether or not the covariates and the spatial
field interact with one another.

  logit_(Pi_i) = Intercept + clay_percent_30cm.std_i + u_i, where u_i is
  the spatially correlated random effect.
  
  Note: this model does not include any other covariates besides percentage
  of clay, as we think this perhaps has the most spatial dependency.

The logit link function is very similar to those written above:

           exp(Intercept + log1p(age.years)_i + u_i)  
  Pi_i = --------------------------------------------
         1 + exp(Intercept + log1p(age.years)_i + u_i)  

Keep in mind that the smoother we included for weighted distance to fire may
not work well if the covariate values are clumped (i.e., if we don't have consistent
sampling across the sampling area, which we don't), so we should define the
RW1 or RW2 smoothers on a set of knots and interpolate the values of the 
smoother at intermediate values.

#### Subsection 2.2: Smoothers and Priors

*From the Highland Stats course:*

The three main approaches for implementing smoothers in INLA are:
 1. Random walk of order 1 (RW1).
 2. Random walk of order 2 (RW2).
 3. One-dimensional SPDE approach. 

Other options exist, but these three are the most commonly used. 

The RW1 model assumes that successive values of the smoother differ only 
slightly and takes into account only "primary" neighboring information:

 f(x_t) - f(x_{t-1})  ~ N(0, sigma_rw^2)


The RW2 smoother takes into account primary and secondary neighbors, so 
there is essentially more incorporation of "surrounding" information. The
RW2 smoother is defined as:
 
 f(x_t) - 2f(x_{t-1}) + f(x_{t-2})  ~ N(0, sigma_rw^2)
  
For a logit link function, you wouldn't expect sigma_rw to be much larger 
than 2-ish (anything above that is quite large on a logit scale). Hence, the 
following prior for sigma_rw is a sensible choice:

 P(sigma_rw > 2) = 0.05.
 
 Here, we'll also set our priors for the Matérn correlation for the spatial
 random field (SRF). We do that for both the "range" and the "sigma" parameters
 associated with the SRF. You can use the info from earlier, when we created
 the mesh, to help inform your priors. We said that anything small-scale in this
 system is around 50km, and set our mesh resolution to 15km, so our prior for range
 should be a little larger than that.
 We'll define our PC priors as:
 
  P(Range < 25km) = 0.95, the probability that the range is < 25km is 95%
  P(sigma > 2) = 0.05, the probability that the sigma is >2 is 5%

----------------------------------------------
If you fit the model with RW1 and RW2 terms, then most of the time the 
computation time will be rather excessive. In `mgcv`, this happens because random 
intercepts are modelled via a smoother function. Here, the issue arises from the 
large number of unique covariate values - for example, weighted distance to 
fire has 88 unique values, which isn't huge, but isn't small, either.

When there are many unique covariate values, the model tries to estimate 
a separate effect for each one, which becomes very slow to compute. To avoid 
this, we can use the same trick as the `bam()` function with its `discrete` option, 
by working with a coarser grid of values instead (basically a shortcut). This is 
where we set the number of knots below.

The code below constructs one-dimensional meshes (sets of equally spaced knots) 
for the weighted distance to fire. The RW smoothers are defined on these mesh 
nodes, not directly at every observed covariate value. INLA estimates the latent 
RW effect at the 20 mesh nodes (the “knots”), with a random walk prior linking 
the neighboring nodes. For each observation, the `bru_mapper()` object then 
interpolates between the two nearest knots, so the smoother value at that covariate 
is a weighted combination of the latent values at the surrounding mesh nodes. 
In other words, we have only 20 latent RW coefficients per covariate, but we 
obtain a smooth value for every observation by interpolation. This reduces the 
dimension of the latent field from “the number of unique covariate values” to 
20 per covariate, making computation much faster, while still providing a 
smooth effect across the full covariate range. 

```{r smoothers-and-priors}
#' Step 1: Setting priors -----
#' Setting the RW penalized complexity (PC) prior for the RW:
PC.rw <- list(prec = list(prior = "pc.prec",
                          param = c(2, 0.05)))
#' This says we think there is only a 5% chance that the sigma_RW is > 2.

#' Setting the PC priors for range and sigma of the SRF. This uses the mesh 
#' we made earlier and the `INLA` package.
Matern <- inla.spde2.pcmatern(mesh1,
                              prior.range = c(25, 0.95),
                              prior.sigma = c(2, 0.05))

#' Step 2: Defining the smoothers -----
#' Defining the one-dimensional mesh for the smoother
#' The `fm_mesh_1d()` function is from the `fmesher` package, and the `bru_mapper()`
#' function is from the `inlabru` package.
mesh.fire <- fm_mesh_1d(seq(min(schisto.complete.sf$weighted_dist_to_fire),
                             max(schisto.complete.sf$weighted_dist_to_fire),
                             length.out = 25)) #' where you set your number of knots
mapper.fire <- bru_mapper(mesh.fire, indexed = FALSE)
#' setting indexed = FALSE ensures that the mapper uses the actual values, rather
#' than indexed (smoother) values, to compute the RW interpolation between knots.
#' Note that the standardized covariates are NOT used in the RW function.

#' Repeat for the other environmental covariates. We can set different knots
#' for each of these 1d meshes, if we'd like.
mesh.twi <- fm_mesh_1d(seq(min(schisto.complete.sf$topo_wetness_index),
                             max(schisto.complete.sf$topo_wetness_index),
                             length.out = 25)) #' where you set your number of knots
mapper.twi <- bru_mapper(mesh.twi, indexed = FALSE)

mesh.msi <- fm_mesh_1d(seq(min(schisto.complete.sf$jan_apr_23_msi),
                             max(schisto.complete.sf$jan_apr_23_msi),
                             length.out = 25)) #' where you set your number of knots
mapper.msi <- bru_mapper(mesh.msi, indexed = FALSE)

#' The smoothers for soil and EVI have been deleted, but they can always be
#' included here again if necessary.
```

#### Subsection 2.3: Defining model components and the likelihood

We'll first fit the model *without* spatial dependency and use that model as 
the reference model.

In `inlabru()`, there are always 3 steps before fitting a model:
 1. Defining the components - what is going into the model (the "suitcase")
 2. Defining the likelihood/model formula
 3. Defining output options (decide what numerical output you'd like)

Keep in mind that `inlabru()` treats factors as random effects in the output.
Also note that when defining components, you can name whatever you'd like *outside*
the parenthesis for that covariate, but what is *inside* the parentheses has to 
match the name in the data frame.

We will be fitting the weighted distance to fire with a "Random Walk" function.
RW order 1 (RW1) penalizes large differences between adjacent values, allowing
the fitted curve to change direction frequently, producing a more "wiggly" 
smoother. This is good for environmental covariates that are locally variable.
RW order 2 (RW2) penalizes large differences in the *slope* between adjacent values
(i.e., it penalizes changes in the first difference between values), resulting
in a smoother, more gradually-changing curve, less "wiggly." This is good for
environmental covariates that change more slowly and continuously.

RW1 is more sensitive to local data sparsity and is prone to overfitting in 
areas with few observations, and since we have larger areas of KNP without 
capture locations, we'll opt for a RW2 for our models.

```{r model-components}
#' Model 1: Bernoulli GAM **without** spatial dependency
#' You can put anything you think you *might* need or want in your components
#' list, but you don't have to include them all in your likelihood.
cmp1 <- ~ Intercept(1) +
  Age(log.age,
      model = "linear") +
  BCS(BCS.Average,
      model = "linear") +
  EVI(jan_apr_23_evi.std,
      model = "linear") +
  soil(clay_percent_30cm.std,
       model = "linear") +
  TWI.rw(topo_wetness_index,
      model = "rw2",
      scale.model = TRUE,
      hyper = PC.rw,
      mapper = mapper.twi,
      constr = TRUE) +
  MSI.rw(jan_apr_23_msi,
      model = "rw2",
      scale.model = TRUE,
      hyper = PC.rw,
      mapper = mapper.msi,
      constr = TRUE) +
  fire.rw(weighted_dist_to_fire,
          model = "rw2",
          scale.model = TRUE,
          hyper = PC.rw,
          mapper = mapper.fire,
          constr = TRUE)
#' Again, note that the standardized version of weighted distance to fire is NOT
#' used in the smoother, rather the raw data.
#' If including a factor, you would write that as:
#' fcap.group(fcapgroup,
#'            model = "factor_contrast")

#' Model 2: Bernoulli GAM **with** spatial dependency
#' This is copy-paste from above, just with the SRF added at the end
#' Note that these same components will be used in Model 3, so no need to re-define
#' them.
cmp2 <- ~ Intercept(1) +
  Age(log.age,
      model = "linear") +
  BCS(BCS.Average,
      model = "linear") +
  EVI(jan_apr_23_evi.std,
      model = "linear") +
  soil(clay_percent_30cm.std,
       model = "linear") +
  TWI.rw(topo_wetness_index,
      model = "rw2",
      scale.model = TRUE,
      hyper = PC.rw,
      mapper = mapper.twi,
      constr = TRUE) +
  MSI.rw(jan_apr_23_msi,
      model = "rw2",
      scale.model = TRUE,
      hyper = PC.rw,
      mapper = mapper.msi,
      constr = TRUE) +
  fire.rw(weighted_dist_to_fire,
          model = "rw2",
          scale.model = TRUE,
          hyper = PC.rw,
          mapper = mapper.fire,
          constr = TRUE) + 
  SRF(geometry,
      model = Matern)
```

Now we'll define our likelihood, which is the same as specifying the model formula.
Here you can pick any or all of the components we defined in the previous step.
Just keep in mind that here you need to use the covariate names that we defined
above *outside* of the parentheses.

```{r model-likelihood}
#' Model 1: Bernoulli GAM **without** spatial dependency
#' Note that the `bru_obs()` command comes from the `inlabru` package
lik1 <- bru_obs(result01 ~ Intercept + Age + BCS + EVI + soil + TWI.rw + MSI.rw +  
                  fire.rw, 
                family = "binomial",
                data = schisto.complete.sf)

#' Model 2: Bernoulli GAM **with** spatial dependency
lik2 <- bru_obs(result01 ~ Intercept + Age + BCS + EVI + soil + TWI.rw + MSI.rw + 
                  fire.rw + SRF,
                family = "binomial",
                data = schisto.complete.sf)

#' Model 3: Bernoulli GAM with interactions between SRF and covariates
#' The first time running this (3.5.26) used soil + SRF, but subsequent run throughs
#' use Age + SRF
lik3 <- bru_obs(result01 ~ Intercept + Age + SRF,
                family = "binomial",
                data = schisto.complete.sf)
```

Now quickly define the output options for the models, which will all be the same.
We would like the 95% credible intervals for the covariates, the DIC, and the WAIC
for the models. Specifying this here means you won't get all the extra output
info you don't want.

```{r output-options}
#' This will be the same across models
Options <- list(quantiles =  c(0.025, 0.975),
                control.compute = list(dic = TRUE, waic = TRUE),
                control.inla = list(strategy = "gaussian")) #' this is where you can specify
                #'                                            inla's integration strategy
```

### Section 3: Fitting the Models -----

Here we'll fit the models and visualize initial results, before doing any kind
of sensitivity analysis.

```{r fitting-model-1}
#' Model 1: Bernoulli GAM **without** spatial dependency

#' Step 1: Fit the model
mod1 <- bru(components = cmp1, lik1, options = Options)

#' Step 2: Look at initial results
mod1$summary.fixed
mod1$summary.hyperpar
#' for random effects, you would run the command:
#' mod1$summary.random$...

#' Step 3: Plot the smoother
#' First, create a grid of distance to fire values between
#' the observed min and max values.
fire.data <- data.frame(weighted_dist_to_fire = seq(from = min(schisto.complete.sf$weighted_dist_to_fire),
                                        to = max(schisto.complete.sf$weighted_dist_to_fire),
                                        length = 100))

#' Now, give this data to the `predict()` function to get outputs for the smoother
pred.fire1 <- predict(mod1, #' tell it which model to use
                     newdata = fire.data, #' tell it what data to use in its prediction
                     formula = ~fire.rw, #' tell it what smoother you used
                     n.samples = 1000) #' tell it how many predictions to generate

#' And plot
plot_fire1 <- ggplot(data = pred.fire1,
                    aes(x = weighted_dist_to_fire,
                        y = mean)) +
              geom_line() +
              geom_hline(yintercept = 0, lty = 2) +
              geom_ribbon(aes(ymax = q0.975,
                              ymin = q0.025),
                          alpha = 0.2) +
  xlab("Weighted Distance to Fire") + ylab("f(Weighted Distance to Fire)") +
  ggtitle("Model 1") +
  theme(text = element_text(size = 12))
plot_fire1
#' What does this mean biologically?
#' Remember that the weighted distance takes into account fire recency, more so than
#' actual distance, and that higher values correspond to more recent fires (and 
#' close proximity to them).
#' Consider flipping the x axis?
#' The RW2 seems to be a bit smoother than the RW1 output
#' Looks similar in terms of where it doesn't cross 0 to other models.

#' Repeat for the other smoothers
#' TWI
twi.data <- data.frame(topo_wetness_index = seq(from = min(schisto.complete.sf$topo_wetness_index),
                                        to = max(schisto.complete.sf$topo_wetness_index),
                                        length = 100))

pred.twi1 <- predict(mod1, 
                     newdata = twi.data, 
                     formula = ~TWI.rw, 
                     n.samples = 1000) 

plot_twi1 <- ggplot(data = pred.twi1,
                    aes(x = topo_wetness_index,
                        y = mean)) +
              geom_line() +
              geom_hline(yintercept = 0, lty = 2) +
              geom_ribbon(aes(ymax = q0.975,
                              ymin = q0.025),
                          alpha = 0.2) +
  xlab("Topo Wetness Index") + ylab("f(Topo Wetness Index)") +
  ggtitle("Model 1") +
  theme(text = element_text(size = 12))
plot_twi1
#' Non-linear, but closer to linear than I'd expect. Some areas a bit close to
#' not overlapping with 0.

#' Repeat for MSI
msi.data <- data.frame(jan_apr_23_msi = seq(from = min(schisto.complete.sf$jan_apr_23_msi),
                                        to = max(schisto.complete.sf$jan_apr_23_msi),
                                        length = 100))

pred.msi1 <- predict(mod1, 
                     newdata = msi.data, 
                     formula = ~MSI.rw, 
                     n.samples = 1000) 

plot_msi1 <- ggplot(data = pred.msi1,
                    aes(x = jan_apr_23_msi,
                        y = mean)) +
              geom_line() +
              geom_hline(yintercept = 0, lty = 2) +
              geom_ribbon(aes(ymax = q0.975,
                              ymin = q0.025),
                          alpha = 0.2) +
  xlab("Wet Season MSI") + ylab("f(Wet Season MSI)") +
  ggtitle("Model 1") +
  theme(text = element_text(size = 12))
plot_msi1
#' Also non-linear, but not super close to avoiding overlap with 0
```

Now we'll run the same model as above, but with an SRF.

```{r fittint-model-2}
#' Model 2: Bernoulli GAM **with** spatial dependency

#' Step 1: Fit the model
mod2 <- bru(components = cmp2, lik2, options = Options)

#' Step 2: Look at initial results
mod2$summary.fixed
mod2$summary.hyperpar
#' for random effects, you would run the command:
#' mod1$summary.random$...

#' Step 3: Plot the smoother
#' First, create a grid of distance to fire values between
#' the observed min and max values.
fire.data <- data.frame(weighted_dist_to_fire = seq(from = min(schisto.complete.sf$weighted_dist_to_fire),
                                        to = max(schisto.complete.sf$weighted_dist_to_fire),
                                        length = 100))

#' Now, give this data to the `predict()` function to get outputs for the smoother
pred.fire2 <- predict(mod2, #' tell it which model to use
                     newdata = fire.data, #' tell it what data to use in its prediction
                     formula = ~fire.rw, #' tell it what smoother you used
                     n.samples = 1000) #' tell it how many predictions to generate

#' And plot
plot_fire2 <- ggplot(data = pred.fire2,
                    aes(x = weighted_dist_to_fire,
                        y = mean)) +
              geom_line() +
              geom_hline(yintercept = 0, lty = 2) +
              geom_ribbon(aes(ymax = q0.975,
                              ymin = q0.025),
                          alpha = 0.2) +
  xlab("Weighted Distance to Fire") + ylab("f(Weighted Distance to Fire)") +
  ggtitle("Model 2") +
  theme(text = element_text(size = 12))
plot_fire2
#' Including the SRF reduced the effect of the smoother for weighted distance
#' to fire, which makes some sense, but it still doesn't cross 0 at lower x-values.

#' Put the two smoother outputs together in one figure for ease
fire_plots <- plot_grid(plot_fire1, plot_fire2, ncol = 2)
fire_plots

#' TWI
twi.data <- data.frame(topo_wetness_index = seq(from = min(schisto.complete.sf$topo_wetness_index),
                                        to = max(schisto.complete.sf$topo_wetness_index),
                                        length = 100))

pred.twi2 <- predict(mod2, 
                     newdata = twi.data, 
                     formula = ~TWI.rw, 
                     n.samples = 1000) 

plot_twi2 <- ggplot(data = pred.twi2,
                    aes(x = topo_wetness_index,
                        y = mean)) +
              geom_line() +
              geom_hline(yintercept = 0, lty = 2) +
              geom_ribbon(aes(ymax = q0.975,
                              ymin = q0.025),
                          alpha = 0.2) +
  xlab("Topo Wetness Index") + ylab("f(Topo Wetness Index)") +
  ggtitle("Model 2") +
  theme(text = element_text(size = 12))
plot_twi2

twi_plots <- plot_grid(plot_twi1, plot_twi2, ncol = 2)
twi_plots
#' Including the SRF changed the relationship at higher x-axis values, but the
#' line is overall much more flat compared to model 1.

#' Repeat for MSI
msi.data <- data.frame(jan_apr_23_msi = seq(from = min(schisto.complete.sf$jan_apr_23_msi),
                                        to = max(schisto.complete.sf$jan_apr_23_msi),
                                        length = 100))

pred.msi2 <- predict(mod2, 
                     newdata = msi.data, 
                     formula = ~MSI.rw, 
                     n.samples = 1000) 

plot_msi2 <- ggplot(data = pred.msi2,
                    aes(x = jan_apr_23_msi,
                        y = mean)) +
              geom_line() +
              geom_hline(yintercept = 0, lty = 2) +
              geom_ribbon(aes(ymax = q0.975,
                              ymin = q0.025),
                          alpha = 0.2) +
  xlab("Wet Season MSI") + ylab("f(Wet Season MSI)") +
  ggtitle("Model 2") +
  theme(text = element_text(size = 12))
plot_msi2

msi_plots <- plot_grid(plot_msi1, plot_msi2, ncol = 2)
msi_plots
#' Including the SRF changed the relationship at the extremes of the x-axis, 
#' but it really flattened the line a lot.

```

And now we'll run model 3: log1p(age) and SRF only

```{r fitting-model-3}
#' Model 3: Bernoulli GAM with an interaction between SRF and covariates

#' Step 1: Fit the model
mod3 <- bru(components = cmp2, lik3, options = Options)

#' Step 2: Look at initial results
mod3$summary.fixed
mod3$summary.hyperpar
#' for random effects, you would run the command:
#' mod1$summary.random$...

#' There is no smoother here to plot, but the range for the SRF increases when 
#' you take away the other covariates. The sigma stays much the same.
```

### Section 4: Model Validation -----

Here we'll look at scaled quantile residuals, uniformity in residuals, and check
spatial dependency for each model.

```{r validation-mod1}
#' Model 1: Bernoulli GAM **without** spatial dependency

#' Step 1: Get scaled quantile residuals
#' To do this, we first need to get the fitted values from the model
N1 <- nrow(schisto.complete.sf)
schisto.complete.sf$Pi.mod1 <- mod1$summary.fitted.values[1:N1, "mean"]

#' Start a loop simulating 1000 Bernoulli distributed observations for each 
#' given fitted value, Pi.mod1. We'll use the `rbinom()` function from the `stats`
#' package.
Ysim1 <- matrix(nrow = N1, ncol = 1000)
for (i in 1:1000){
  Ysim1[,i] <- rbinom(n = N1,
                      size = 1,
                      prob = schisto.complete.sf$Pi.mod1)
}

#' Now pass this simulated data through the `DHARMa` package, which will calculate
#' scaled quantile residuals for you.
E1sqr <- createDHARMa(simulatedResponse = Ysim1,
                      observedResponse = schisto.complete.sf$result01,
                      fittedPredictedResponse = schisto.complete.sf$Pi.mod1,
                      integerResponse = TRUE)

#' If you were running a Poisson model, here's where you would check for
#' overdispersion.

#' Step 2: Check for residual uniformity
#' The `DHARMa` package will also do this for you!
par(mfrom = c(1,1), mar = c(5,5,2,2))
plotQQunif(E1sqr,
           testUniformity = TRUE,
           testOutliers = TRUE,
           testDispersion = FALSE)
#' Seems ok!

#' Step 3: Plot the residuals vs the covariates using `DHARMa`
plotResiduals(E1sqr, form = schisto.complete.sf$log.age) #' Ok
plotResiduals(E1sqr, form = schisto.complete.sf$BCS.Average) #' Ok
plotResiduals(E1sqr, form = schisto.complete.sf$topo_wetness_index) #' Ok
plotResiduals(E1sqr, form = schisto.complete.sf$clay_percent_30cm) #' Ok
plotResiduals(E1sqr, form = schisto.complete.sf$jan_apr_23_evi) #' Ok
plotResiduals(E1sqr, form = schisto.complete.sf$jan_apr_23_msi) #' Ok
plotResiduals(E1sqr, form = schisto.complete.sf$weighted_dist_to_fire) #' Ok

#' Step 4: Check for spatial dependency
#' There are 3 ways we can do this:
#'  1. Plot residuals vs spatial locations and look for patterns
#'  2. Apply a Moran's I test on the residuals
#'  3. Make a variogram of the residuals
#' We'll do all of these just for fun.

#' Option 1: Plot residuals vs spatial locations
#' The `residuals()` function comes from the `stats` package.
schisto.complete.sf$E1 <- residuals(E1sqr)

#' Define the color and point size for the plot
#' Remember that scaled quantile residuals are centered on 0.5, while Pearson 
#' residuals are centered around 0.
schisto.complete.sf$MyCol <- ifelse(schisto.complete.sf$E1 >= 0.5, "red", "blue")
schisto.complete.sf$MySize <- rescale(abs(schisto.complete.sf$E1), to = c(0,3))

#' Plot it
mod1residspat <- ggplot(data = kruger.km) +
  geom_sf(fill = "transparent") +
  geom_sf(data = schisto.complete.sf,
          aes(col = MyCol,
              size = MySize),
          alpha = 0.3) +
  scale_color_identity() +
  scale_size_continuous(range = c(1,3)) +
  theme_minimal() +
  theme(legend.position = "none") +
  guides(fill = guide_legend(title = NULL)) +
  labs(title = "Residuals Mod 1")
mod1residspat

#' Hmm, maybe a bit of a spatial pattern in the southeast?

#' Option 2: Moran's I test
#' `DHARMa` will do this for you; however, the function doesn't do well with
#' observations that have the exact same coordinates, so we'll add a few
#' cms to each location.
testSpatialAutocorrelation(E1sqr, 
                           x = schisto.complete.sf$Xkm + rnorm(N1, mean = 0, sd = 0.0001), 
                           y = schisto.complete.sf$Ykm + rnorm(N1, mean = 0, sd = 0.0001))
#' There could be some minor spatial dependency in the residuals.

#' Option 3: Make a variogram
#' First, make a dataframe with the residuals and capture locations
mod1data <- data.frame(E1 = schisto.complete.sf$E1,
                       Xkm = schisto.complete.sf$Xkm,
                       Ykm = schisto.complete.sf$Ykm)

#' Convert to an `sf` object
mod1data_sf <- st_as_sf(x = mod1data,
                        coords = c("Xkm", "Ykm"),
                        crs = NA) #' CRS doesn't matter for this, so no need to specify

#' Call the `variogram()` function from the package `gstat`
V1 <- variogram(E1 ~1,
                data = mod1data_sf,
                cutoff = 100,
                cressie = TRUE)

#' And plot
variogrammod1 <- ggplot(data = V1, aes(x = dist, y = gamma)) +
                 geom_point() +
                 geom_smooth(se = FALSE, span = 1) +
                 labs(x = "Distance (km)", y = "Semi-variogram Mod 1") +
                 coord_cartesian(ylim = c(0.083, 0.093)) +
                 theme(text = element_text(size = 15))
variogrammod1
#' Peak spatial correlation is up to ~37km; however, it's important to note that
#' the spatial correlation never falls below 0.86!
```

Model 1 conclusions: 
- We may have a slight spatial effect?
- There is a non-linear effect of weighted distance to fire, TWI, MSI, but perhaps
mostly just with fire.
- Interestingly, BCS comes out as close important here in model 2 (positive effect),
and including just age + SRF flips the age effect from negative to positive 
(although still not important).

```{r validation-mod2}
#' Model 2: Bernoulli GAM **with** spatial dependency

#' Step 1: Get scaled quantile residuals
#' To do this, we first need to get the fitted values from the model
N2 <- nrow(schisto.complete.sf)
schisto.complete.sf$Pi.mod2 <- mod2$summary.fitted.values[1:N2, "mean"]

#' Start a loop simulating 1000 Bernoulli distributed observations for each 
#' given fitted value, Pi.mod1. We'll use the `rbinom()` function from the `stats`
#' package.
Ysim2 <- matrix(nrow = N2, ncol = 1000)
for (i in 1:1000){
  Ysim2[,i] <- rbinom(n = N2,
                      size = 1,
                      prob = schisto.complete.sf$Pi.mod2)
}

#' Now pass this simulated data through the `DHARMa` package, which will calculate
#' scaled quantile residuals for you.
E2sqr <- createDHARMa(simulatedResponse = Ysim2,
                      observedResponse = schisto.complete.sf$result01,
                      fittedPredictedResponse = schisto.complete.sf$Pi.mod2,
                      integerResponse = TRUE)

#' If you were running a Poisson model, here's where you would check for
#' overdispersion.

#' Step 2: Check for residual uniformity
#' The `DHARMa` package will also do this for you!
par(mfrom = c(1,1), mar = c(5,5,2,2))
plotQQunif(E2sqr,
           testUniformity = TRUE,
           testOutliers = TRUE,
           testDispersion = FALSE)
#' Seems ok!

#' Step 3: Plot the residuals vs the covariates using `DHARMa`
plotResiduals(E2sqr, form = schisto.complete.sf$log.age) #' Ok
plotResiduals(E2sqr, form = schisto.complete.sf$BCS.Average) #' Ok
plotResiduals(E2sqr, form = schisto.complete.sf$topo_wetness_index) #' Ok
plotResiduals(E2sqr, form = schisto.complete.sf$clay_percent_30cm) #' Ok
plotResiduals(E2sqr, form = schisto.complete.sf$jan_apr_23_evi) #' Ok
plotResiduals(E2sqr, form = schisto.complete.sf$jan_apr_23_msi) #' Ok
plotResiduals(E2sqr, form = schisto.complete.sf$weighted_dist_to_fire) #' Ok

#' Step 4: Check for spatial dependency
#' There are 3 ways we can do this:
#'  1. Plot residuals vs spatial locations and look for patterns
#'  2. Apply a Moran's I test on the residuals
#'  3. Make a variogram of the residuals
#' We'll do all of these just for fun.

#' Option 1: Plot residuals vs spatial locations
#' The `residuals()` function comes from the `stats` package.
schisto.complete.sf$E2 <- residuals(E2sqr)

#' Define the color and point size for the plot
#' Remember that scaled quantile residuals are centered on 0.5, while Pearson 
#' residuals are centered around 0.
schisto.complete.sf$MyCol2 <- ifelse(schisto.complete.sf$E2 >= 0.5, "red", "blue")
schisto.complete.sf$MySize2 <- rescale(abs(schisto.complete.sf$E2), to = c(0,3))

#' Plot it
mod2residspat <- ggplot(data = kruger.km) +
  geom_sf(fill = "transparent") +
  geom_sf(data = schisto.complete.sf,
          aes(col = MyCol2,
              size = MySize2),
          alpha = 0.3) +
  scale_color_identity() +
  scale_size_continuous(range = c(1,3)) +
  theme_minimal() +
  theme(legend.position = "none") +
  guides(fill = guide_legend(title = NULL)) +
  labs(title = "Residuals Mod 2")
mod2residspat
#' Hmm, maybe resolved some of that spatial pattern in the SE?

#' Option 2: Moran's I test
#' `DHARMa` will do this for you; however, the function doesn't do well with
#' observations that have the exact same coordinates, so we'll add a few
#' cms to each location.
testSpatialAutocorrelation(E2sqr, 
                           x = schisto.complete.sf$Xkm + rnorm(N2, mean = 0, sd = 0.0001), 
                           y = schisto.complete.sf$Ykm + rnorm(N2, mean = 0, sd = 0.0001))
#' There could be some minor spatial dependency in the residuals?

#' Option 3: Make a variogram
#' First, make a dataframe with the residuals and capture locations
mod2data <- data.frame(E2 = schisto.complete.sf$E2,
                       Xkm = schisto.complete.sf$Xkm,
                       Ykm = schisto.complete.sf$Ykm)

#' Convert to an `sf` object
mod2data_sf <- st_as_sf(x = mod2data,
                        coords = c("Xkm", "Ykm"),
                        crs = NA) #' CRS doesn't matter for this, so no need to specify

#' Call the `variogram()` function from the package `gstat`
V2 <- variogram(E2 ~1,
                data = mod2data_sf,
                cutoff = 100,
                cressie = TRUE)

#' And plot
variogrammod2 <- ggplot(data = V2, aes(x = dist, y = gamma)) +
                 geom_point() +
                 geom_smooth(se = FALSE, span = 1) +
                 labs(x = "Distance (km)", y = "Semi-variogram Mod 2") +
                 coord_cartesian(ylim = c(0.083, 0.093)) +
                 theme(text = element_text(size = 15))
variogrammod2
#' Now we've maybe done some inverse correlation?
#' #' It's important to note that nothing drops below spatial correlation of 
#' ~0.88.
```

Model 2 conclusions:
- Including the SRF reduced the effect of the weighted distance to fire smoother
- Weighted distance to fire still has a non-linear effect
- Possible non-linear effects of TWI and MSI, but not nearly as strong as weighted
distance to fire.
- We may have introduced inverse spatial correlation?
- Other covariates still include 0 in their CIs

```{r validation-mod3}
#' Model 3: Bernoulli GLM with SRF and one covariate

#' Step 1: Get scaled quantile residuals
#' To do this, we first need to get the fitted values from the model
N3 <- nrow(schisto.complete.sf)
schisto.complete.sf$Pi.mod3 <- mod3$summary.fitted.values[1:N3, "mean"]

#' Start a loop simulating 1000 Bernoulli distributed observations for each 
#' given fitted value, Pi.mod1. We'll use the `rbinom()` function from the `stats`
#' package.
Ysim3 <- matrix(nrow = N3, ncol = 1000)
for (i in 1:1000){
  Ysim3[,i] <- rbinom(n = N3,
                      size = 1,
                      prob = schisto.complete.sf$Pi.mod3)
}

#' Now pass this simulated data through the `DHARMa` package, which will calculate
#' scaled quantile residuals for you.
E3sqr <- createDHARMa(simulatedResponse = Ysim3,
                      observedResponse = schisto.complete.sf$result01,
                      fittedPredictedResponse = schisto.complete.sf$Pi.mod3,
                      integerResponse = TRUE)

#' If you were running a Poisson model, here's where you would check for
#' overdispersion.

#' Step 2: Check for residual uniformity
#' The `DHARMa` package will also do this for you!
par(mfrom = c(1,1), mar = c(5,5,2,2))
plotQQunif(E3sqr,
           testUniformity = TRUE,
           testOutliers = TRUE,
           testDispersion = FALSE)
#' Seems ok!

#' Step 3: Plot the residuals vs the covariates using `DHARMa`
plotResiduals(E3sqr, form = schisto.complete.sf$log.age) #' Ok

#' Step 4: Check for spatial dependency
#' There are 3 ways we can do this:
#'  1. Plot residuals vs spatial locations and look for patterns
#'  2. Apply a Moran's I test on the residuals
#'  3. Make a variogram of the residuals
#' We'll do all of these just for fun.

#' Option 1: Plot residuals vs spatial locations
#' The `residuals()` function comes from the `stats` package.
schisto.complete.sf$E3 <- residuals(E3sqr)

#' Define the color and point size for the plot
#' Remember that scaled quantile residuals are centered on 0.5, while Pearson 
#' residuals are centered around 0.
schisto.complete.sf$MyCol3 <- ifelse(schisto.complete.sf$E3 >= 0.5, "red", "blue")
schisto.complete.sf$MySize3 <- rescale(abs(schisto.complete.sf$E3), to = c(0,3))

#' Plot it
mod3residspat <- ggplot(data = kruger.km) +
  geom_sf(fill = "transparent") +
  geom_sf(data = schisto.complete.sf,
          aes(col = MyCol3,
              size = MySize3),
          alpha = 0.3) +
  scale_color_identity() +
  scale_size_continuous(range = c(1,3)) +
  theme_minimal() +
  theme(legend.position = "none") +
  guides(fill = guide_legend(title = NULL)) +
  labs(title = "Residuals Mod 3")
mod3residspat
#' Hmm, not sure - looks really similar to the output of model 2.

#' Option 2: Moran's I test
#' `DHARMa` will do this for you; however, the function doesn't do well with
#' observations that have the exact same coordinates, so we'll add a few
#' cms to each location.
testSpatialAutocorrelation(E3sqr, 
                           x = schisto.complete.sf$Xkm + rnorm(N3, mean = 0, sd = 0.0001), 
                           y = schisto.complete.sf$Ykm + rnorm(N3, mean = 0, sd = 0.0001))
#' There could be some minor spatial dependency in the residuals?

#' Option 3: Make a variogram
#' First, make a dataframe with the residuals and capture locations
mod3data <- data.frame(E3 = schisto.complete.sf$E3,
                       Xkm = schisto.complete.sf$Xkm,
                       Ykm = schisto.complete.sf$Ykm)

#' Convert to an `sf` object
mod3data_sf <- st_as_sf(x = mod3data,
                        coords = c("Xkm", "Ykm"),
                        crs = NA) #' CRS doesn't matter for this, so no need to specify

#' Call the `variogram()` function from the package `gstat`
V3 <- variogram(E3 ~1,
                data = mod3data_sf,
                cutoff = 100,
                cressie = TRUE)

#' And plot
variogrammod3 <- ggplot(data = V3, aes(x = dist, y = gamma)) +
  geom_point() +
  geom_smooth(se = FALSE, span = 1) +
  labs(x = "Distance (km)", y = "Semi-variogram Mod 3") +
  coord_cartesian(ylim = c(0.083, 0.093)) +
  theme(text = element_text(size = 15))
variogrammod3
# This looks similar to the output of model 2.
#' It's important to note that nothing drops below spatial correlation of 
#' ~0.88.
```

We'll combine the residual spatial output figures in to one multipanel figure
for ease of viewing.

```{r combo-validation-figures}
#' Combo spatial variation in residuals
spatresid_plots <- plot_grid(mod1residspat, mod2residspat, mod3residspat, ncol = 3)
spatresid_plots

#' Combo variogram plots
variogram_combo_plots <- plot_grid(variogrammod1, variogrammod2, variogrammod3, ncol = 3)
variogram_combo_plots
```

It seems as though adding in the environmental covariates does at least decrease
the strength of the negative spatial correlation (mod 2 is an improvement in that
regard compared to mod 3).

### Section 5: Model Comparison -----

Now we'll compare all 3 models using DIC and WAIC (the Bayesian equivalent of AIC).

```{r model-comparison}
DICs  <- c(mod1$dic$dic, mod2$dic$dic, mod3$dic$dic)
WAICs <- c(mod1$waic$waic, mod2$waic$waic, mod3$waic$waic)
edf.dic  <- c(mod1$dic$p.eff, mod2$dic$p.eff, mod3$dic$p.eff)
waic.dic <- c(mod1$waic$p.eff, mod2$waic$p.eff, mod3$waic$p.eff)

Out   <- data.frame(Models = c("Bernoulli GAM",
                               "Bernoulli GAM + SRF",
                               "Bernoulli GAM + age + SRF"),
                    DIC = DICs,
                    edf.dic = edf.dic,
                    WAIC = WAICs,
                    waic.dic)
colnames(Out) <- c(" GAM model", 
                   "       DIC",
                   "  ENP.dic",
                   "      WAIC",
                   " ENP.waic")
Out
```

**From the first runthrough**
Interestingly, the model with no covariates (apart from the % clay in the soil) 
has the lowest DIC/WAIC values, but the difference isn't all that much. I think 
likely the covariates and the SRF are explaining the same thing, so including 
both in the model isn't that helpful.

**From the second runthrough**
Increasing mesh resolution and setting better priors decreased the gap between 
the WAIC/DIC values for models 2 and 3. So technically, although the GAM with
age + SRF has the lowest WAIC, it isn't all that different from the model with
the environmental covariates AND the SRF.

**From the third runthrough**
Modeling the other environmental covariates with a RW smoother really improved
the model, such that it actually has a *lower* WAIC (although not significantly)
compared to model 3. This is a change, though, from previous model runs. Model 3
does have the lowest number of effective parameters still, though.
It does seem like EVI might have more of a linear relationship, so perhaps it 
doesn't necessarily need to be modeled as a RW2, or maybe it would be better
modeled as a RW1?

**From the fourth runthrough**
Modeling with fire, TWI, and MSI  as RW2 smoothers does have a lower WAIC than
model 3, but it's still within 5, so there isn't a significant difference.
Both models 2 and 3 have lower WAICs compared to model 1, but model 2 does have
the highest number of effective parameters. We still need to check if this
overfits the model like the third runthrough did, and it still may be more 
appropriate to model TWI and MSI as linear.

We'll plot the SRFs.

```{r srf-plot}
#' First, we need to convert the boundaries to an sfc object for plotting. The
#' sfc object is just the geometry, without any additional attribute data attached.
#' The inner hull can be used if you only want to plot the SRF for areas that are
#' closely wrapped to your data.
InnerHull_sfc <- fm_as_sfc(inner.hull) 
st_crs(InnerHull_sfc) <- st_crs(schisto.complete.sf) 

#' We'll use the outer hull, so that we have all of KNP covered
OuterHull_sfc <- st_geometry(outer_sf_simple)

NewData <- fm_pixels(mesh1, 
                      dims = c(250, 250),
                      mask = OuterHull_sfc)

#' Plot model 2
Pred <- predict(object = mod2, 
                newdata = NewData, 
                formula = ~ SRF)

#' Some nice colors
ColSc <- function(...) {
  scale_fill_gradientn(
    colours = rev(RColorBrewer::brewer.pal(11, "RdYlBu")),
    limits = range(..., na.rm = TRUE))}

#' Plot the spatial random field.
mod2SRF <- ggplot() +
  gg(Pred, geom = "tile") +
  geom_sf(data = kruger.km,
          fill = "transparent", 
          col = "darkred") +
  ggtitle("Posterior mean SRF Mod 2") +
  labs(x = "Easting (km)", y = "Northing (km)") + 
  ColSc(Pred$mean)
mod2SRF

#' Plot model 3
Pred <- predict(object = mod3, 
                newdata = NewData, 
                formula = ~ SRF)

#' Some nice colors
ColSc <- function(...) {
  scale_fill_gradientn(
    colours = rev(RColorBrewer::brewer.pal(11, "RdYlBu")),
    limits = range(..., na.rm = TRUE))}

#' Plot the spatial random field.
mod3SRF <- ggplot() +
  gg(Pred, geom = "tile") +
  geom_sf(data = kruger.km,
          fill = "transparent", 
          col = "darkred") +
  ggtitle("Posterior mean SRF Mod 3") +
  labs(x = "Easting (km)", y = "Northing (km)") + 
  ColSc(Pred$mean)
mod3SRF

#' Combine the plots
comboSRF <- plot_grid(mod2SRF, mod3SRF, ncol = 2)
comboSRF
```

The figure above shows the posterior mean of the spatial random field. Warm 
colors indicate areas where the spatial field is associated with positive schisto, 
status, and cool colors indicate areas where it is associated with negative 
schisto status, after accounting for covariate effects.

It looks like modeling with the other environmental covariates fit with RW2
smoothers may have overfit the model, and modeling with fire, TWI, and MSI as
RW2s (EVI and soil as linear), still overfits the model.

### Section 6: Imposed Spatial Dependency -----

Here we can look visually at the posterior distribution of the range and the 
imposed spatial dependence in the model. These looked similar for models 2 and 3.

We can then define distances where the correlation decays:

  - strong correlation as correlation between 1 - 0.8.
  - moderate correlation as correlation between 0.8 - 0.5.
  - weak correlation as correlation between 0.5 - 0.1.
  - diminishing correlation as correlation smaller than 0.1.

```{r spatial-dependency}
#' Step 1: Get the posterior distribution of the range
mod2Range <- spde.posterior(mod2, "SRF", what = "range")
mod2rangeplot <- plot(mod2Range)

mod3Range <- spde.posterior(mod3, "SRF", what = "range")
mod3rangeplot <- plot(mod3Range)

comborangeplots <- plot_grid(mod2rangeplot, mod3rangeplot, ncol = 1)
comborangeplots

#' Step 2: Get the imposed spatial dependence for each model
#' Model 2
MaternCor.mod2 <- spde.posterior(mod2, "SRF", what = "matern.correlation")
pSpatDep.mod2 <- plot(MaternCor.mod2) + labs(x = "Distance (km)", y = "Correlation") +
            coord_cartesian(ylim = c(0, 1))
pSpatDep.mod2

#' Model 3
MaternCor.mod3 <- spde.posterior(mod3, "SRF", what = "matern.correlation")
pSpatDep.mod3 <- plot(MaternCor.mod3) + labs(x = "Distance (km)", y = "Correlation") +
            coord_cartesian(ylim = c(0, 1))
pSpatDep.mod3

#' And add code from the Highland Statistics folks to include the distances at 
#' which the correlation hits a given level (0.8, 0.5, and 0.1 thresholds).
gb   <- ggplot_build(pSpatDep.mod2)
dfB  <- gb$data[[1]][, c("x","y")]
dfB  <- dfB[order(dfB$x), ]
dfB2 <- dfB[order(dfB$y), ]  # y ascending, x matched

x_at <- function(level) {
  approx(x = dfB2$y, y = dfB2$x, xout = level, rule = 2)$y
}

thr   <- c(0.8, 0.5, 0.1)
xthr  <- sapply(thr, x_at)
marks <- data.frame(level = thr, x = xthr)

xmax <- max(dfB$x)
p.Matern.mod2 <- pSpatDep.mod2 +
  geom_segment(data = marks, 
               aes(x = 0, xend = x, y = level, yend = level), 
               linetype = "dotted") +
  geom_segment(data = marks, 
               aes(x = x,  xend = x, y = 0, yend = level), 
               linetype = "dotted") +
  geom_point(data = marks, 
             aes(x = x, y = level), 
             size = 1.2) +
  annotate("text", 
           x = xthr, y = 0, 
           label = paste0("~", round(xthr, 1)),
           vjust = 1.4, 
           size = 3) +
  annotate("text", 
           x = 0.97 * xmax, y = 0.90, 
           hjust = 1, 
           size = 3.2,
           label = "Strong: 1–0.8") +
  annotate("text", 
           x = 0.97 * xmax, 
           y = 0.65, 
           hjust = 1, 
           size = 3.2,
           label = "Moderate: 0.8–0.5") +
  annotate("text", 
           x = 0.97 * xmax, 
           y = 0.30, 
           hjust = 1, 
           size = 3.2,
           label = "Weak: 0.5–0.1") +
  annotate("text", 
           x = 0.97 * xmax, 
           y = 0.06, 
           hjust = 1, 
           size = 3.2,
           label = "Diminishing: < 0.1") +
  theme_minimal() +
  ggtitle("Mod 2 Spatial Dependence") +
  theme(text = element_text(size = 15)) 
p.Matern.mod2

#' Repeat for mod 3
gb   <- ggplot_build(pSpatDep.mod3)
dfB  <- gb$data[[1]][, c("x","y")]
dfB  <- dfB[order(dfB$x), ]
dfB2 <- dfB[order(dfB$y), ]  # y ascending, x matched

x_at <- function(level) {
  approx(x = dfB2$y, y = dfB2$x, xout = level, rule = 2)$y
}

thr   <- c(0.8, 0.5, 0.1)
xthr  <- sapply(thr, x_at)
marks <- data.frame(level = thr, x = xthr)

xmax <- max(dfB$x)
p.Matern.mod3 <- pSpatDep.mod3 +
  geom_segment(data = marks, 
               aes(x = 0, xend = x, y = level, yend = level), 
               linetype = "dotted") +
  geom_segment(data = marks, 
               aes(x = x,  xend = x, y = 0, yend = level), 
               linetype = "dotted") +
  geom_point(data = marks, 
             aes(x = x, y = level), 
             size = 1.2) +
  annotate("text", 
           x = xthr, y = 0, 
           label = paste0("~", round(xthr, 1)),
           vjust = 1.4, 
           size = 3) +
  annotate("text", 
           x = 0.97 * xmax, y = 0.90, 
           hjust = 1, 
           size = 3.2,
           label = "Strong: 1–0.8") +
  annotate("text", 
           x = 0.97 * xmax, 
           y = 0.65, 
           hjust = 1, 
           size = 3.2,
           label = "Moderate: 0.8–0.5") +
  annotate("text", 
           x = 0.97 * xmax, 
           y = 0.30, 
           hjust = 1, 
           size = 3.2,
           label = "Weak: 0.5–0.1") +
  annotate("text", 
           x = 0.97 * xmax, 
           y = 0.06, 
           hjust = 1, 
           size = 3.2,
           label = "Diminishing: < 0.1") +
  theme_minimal() +
  ggtitle("Mod 3 Spatial Dependence") +
  theme(text = element_text(size = 15)) 
p.Matern.mod3

#' Combine the two plots
combo_spatdepplots <- plot_grid(p.Matern.mod2, p.Matern.mod3, ncol = 1)
combo_spatdepplots
```

**First runthrough**
This plot tells us that the spatial correlation drops below 0.8 at ~27.8km, below
0.5 at ~62.4km, and below 0.1 at ~159.5km.

**Second runthrough**
This plot tells us that the spatial correlation drops below 0.8 at ~13km, below
0.5 at ~29km, and below 0.1 at ~74.3km - a pretty big difference compared to the
output from the first runthrough with a coarser mesh.

**Third runthrough**
Including the other env covariates as smoothers overfit our model. The spatial
correlation drops below 0.8 at <1km.

**Fourth runthrough**
Including fire, TWI, and MSI as smoothers still leaves us with an overfit model.
The spatial correlation drops below 0.8 at <1km still.

### Overall Conclusions -----

**Last updated 3.17.26**

We should do a sensitivity analysis and see whether we can improve the MeshResolution,
prior of the range, and prior of the sigma_u.
  - Try mesh resolution of 10km?
  - Update the prior to reflect what we know about the range.
  - Extract the sigma we got here and use that to inform what the prior should
  be for sigma_u.
  - Only fire should be included as a smoother. Including other covariates as
  smoothers just overfits our model.

We will also be building a model to look at **schisto burden**, not just disease
status (pos/neg). We can see if any of these environmental covariates are important
for explaining burden on the landscape.