---
title: "Schisto Test INLA-SPDE"
author: "Eberle Yarborough"
date: "`r Sys.Date()`"
output: pdf_document
---

This document is an attempt to test a less-complex INLA-SPDE model of *Schistosoma* spp. burden (represented via CAA titer) to generate a predictive heat map of burden in African buffalo (*Syncerus caffer*) in Kruger National Park, based off of the code found in the `anaplasma_INLA.Rmd`, `anaplasma_INLA_henri.Rmd`, and `limited_INLA.Rmd` files.

Here, we'll try to run the INLA model using buffalo schisto CAA titer as the response variable, with demographic factors, Animal ID, management section, and a single environmental variable (likely elevation) as covariates. When working with CAA titers, animals with values \< 30 (i.e., negative values) must be removed for analysis. These values also tend to follow a negative binomial distribution, which means the model family can be either negative binomial or gamma. For this initial analysis, we will log transform CAA titer and add 1 to all values (since you can't log transform 0).

### General INLA Setup, Packages, Setting wd, etc.

```{r setup}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
  include = TRUE,
	message = FALSE,
	warning = FALSE
)

# On iOS, file paths do not start with "C:/" like on windows, just with a forward slash
knitr::opts_knit$set(root.dir = "/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Schisto/Schisto_Analysis")

# Packages
required_packages <- c("terra", "geodata", "sf", "sp", "terra", "fmesher",
                       "tidyverse", "knitr", 
                       "lme4", "viridis", "Matrix", "tibble",
                       "FNN",
                       "ggeffects", "ggnewscale", "ggtext",
                       "ggfortify", "ggspatial",
                       "stringr", "magrittr",
                       "RColorBrewer", "httr", "scales",
                       "abbreviate", "kableExtra", "texreg", "latex2exp",
                       "patchwork", "readr",
                       "remotes", "devtools", "callr", "CopernicusClimate", "ncdf4"
)#

# Install missing packages and load them
for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, dependencies = TRUE)  # Install package if missing
    library(pkg, character.only = TRUE)         # Load package
  }
}


if(!require(INLA)) install.packages("INLA",
                    repos=c(getOption("repos"),
                    INLA="https://inla.r-inla-download.org/R/stable"),
                    dep=TRUE)
#inla::inla.upgrade() # for the stable version

if(!require(brinla)) remotes::install_github("julianfaraway/brinla")

if(!require(ggregplot)) devtools::install_github("gfalbery/ggregplot") # Installing Greg's package for plotting functions!


## Set working dir
# Set project directory here for easy file paths.
# If you opened this as a project getwd(), should be the project directory

#getwd()

# If you opened this in another way, set the project directory manually below
# There are other "better" ways to do this, but I find them not to be consistent, e.g. the 'here' package

file_path <- file.path("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Schisto/Schisto_Analysis")

# Then set relative file path later on, e.g.:
# file.path(file_path, "shapefiles", "kruger.shp")


# Page parameters
pageheight = 7.64 #11 - 2
pagewidth = 4.7 #8.5 - 2

# plot sizes
textwidth <- 159.2 #mm
textheight <- 236.2 #mm

# Plot themes
par(family = "Arial")

# Set theme
ggTheme <- theme_classic()

# Set plot details
ggChange <- 
theme(#text=element_text(size=8), #change font size of all text
      text = element_text(family = "Helvetica"),
      plot.title=element_text(size=10), #change font size of plot title
      axis.title=element_text(size=10), #change font size of axis titles
      axis.text=element_text(size=8), #change font size of axis text
      legend.title=element_text(size=8),
      legend.text=element_text(size=8) #change font size of legend text
      )#

# Plot settings alternative (called in inla background scripts - did not feel like looking/finding and changing it everywhere)
my_plot_settings <- 
             theme(text = element_text(family = "Helvetica"),
             plot.title = element_text(size = 8),
             plot.subtitle = element_text(size = 8),
             axis.title = element_text(size = 8),
             axis.text = element_text(size = 6),
             legend.text = element_text(size = 6),
             legend.title = element_text(size = 8),
             legend.spacing.y = unit(0.5, 'mm'),
             legend.box.spacing = unit(0.5, "mm"),
             legend.key.size = unit(4, "mm"),
             strip.text = element_text(size = 6))

#gitlab <- "anaplasma_inla"
#github <- "Schisto_Analysis" # this is the same as the parent directory/github directory already
```

These are lines Henri wrote to make sure the github has the necessary directory structure for the code to run appropriately (these have already been created as of 11 Jan 2026):

```{r checking-directory-structure}
# check dir structure - DONE
#if(!file.exists(file.path(file_path, "figures"))) dir.create(file.path(file_path, "figures"))
#if(!file.exists(file.path(file_path, "tables"))) dir.create(file.path(file_path, "tables"))
#if(!file.exists(file.path(file_path, "inla.model"))) dir.create(file.path(file_path, "inla.model"))
#if(!file.exists(file.path(file_path, "inla.model", "dic"))) dir.create(file.path(file_path, "inla.model", "dic"))
#if(!file.exists(file.path(file_path, "inla.model", "logfiles"))) dir.create(file.path(file_path, "inla.model", "logfiles"))
#if(!file.exists(file.path(file_path, "inla.model", "dic", "dic_fieldplots"))) dir.create(file.path(file_path, "inla.model", "dic", "dic_fieldplots"))
```

These are scripts for doing various things that Henri wrote or adapted, as well as some scripts for merging raster data that I created. Some are called later, especially for plotting the spatial fields. They can be run in the background with "background jobs \> script \> run with a copy of global environment \> copy to global environment."

```{r custom-scripts}
# load scripts - use full file path since scrips are in a different directory
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_inla_my_INLARange.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_inla_my_ggField.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_inla_myINLA_extract_2.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_inla_my_ggMesh.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_sommer_myanova.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_ggsave_my_ggsave_plots.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_busywork.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_inla_my_extract_projection_values.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/Anaplasma/gitlab/anaplasma_inla/scripts/rfunction_latex_subfiguregenerator.R")
source("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/KNP Environmental Data/my_raster_load_function.R")
```

Loading in the schisto data, combining with capture locations and environmental layer(s):

```{r loading-data, echo=FALSE}
## Loading the master data frame
base_schisto_df<- read.csv("schisto_bym_data.csv")
# Note: there are 217 observations that are MISSING CAA data - these were below LLOD in the raw data

## Loading environmental data
env_covars <- st_read("env_covar_data_LSTresolution.gpkg")

# Call the safe raster load function to make sure rasters are loaded in cleanly if including new ones, especially when they have duplicate names contained within them (common for EVI/NDVI)

# fall.23.EVI <- safe_raster_load_function("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/KNP Environmental Data/Cleaned Rasters/EVI_MarApr2023_Marginal_Mean.tif", prefix = "evifall23")
# fall.23.EVI <- project(fall.23.EVI, "EPSG:32736")

# Optional: At this point, open the custom `my_raster_merge_function` code if wanting to align new rasters to existing ones in use
# As of 2/19, all are aligned to the coarsest layer of LST day, but if LST is removed from analysis, would need to realign to the new coarsest layer

## Merging the schisto data with the capture locations and environmental variables
# Note: All `raster_data` is on the left of the merge to preserve the spatial data when the files are merged
schisto_spatial_df <- env_covars %>%
  rename(animal.ID = Animal.ID) %>% # making sure the column headers match for merging 
  
  # Now adding the rest of the data
  left_join(base_schisto_df, by = "animal.ID") %>%
  select(
    animal.ID,                                         # Reordering the variables - ID first
    age.years, age.class, Sex, BCS.Average, Section,   # Then demographics
    CAA.Estimate,                                      # Disease variable (response)
    fall.2023.NDVI, spring.2022.NDVI, fall.23.EVI, spring.22.EVI, # Environmental data
    elevation, fall.2023.LST.day.mean, fall.2023.LST.night.mean,
    primary_soil_type, primary_soil_descr, soil_category,
    everything()                                       # Catching anything else I may have missed (geom automatically goes last)
  ) %>%
  
  # Converting management section to numeric since it'll be used later on as a random effect
  dplyr::mutate(Section = as.numeric(as.factor(Section))) %>%
  
  # Converting response variable to numeric, rather than an integer for INLA
  dplyr::mutate(CAA.Estimate = as.numeric(CAA.Estimate))

# Removing a random space from the column name for age
schisto_spatial_df$age.class <- gsub(" ", "", schisto_spatial_df$age.class)

#### NOTE!!! See env_covars_collinearity plot to look at correlation between mean LST day/night, elev, and fall/spring NDVI

# Note: not writing to a df at the moment so variables can be changed out later with ease

# Converting the capture locations back to easting/northing as a matrix, rather than point geometry  - this is how it should be for INLA analysis
# Get UTM coordinates (not lat/long)
buff.locations.utm <- st_coordinates(schisto_spatial_df)  # Already in UTM from st_read

schisto_spatial_df$x.east <- buff.locations.utm[,"X"]  # UTM Easting
schisto_spatial_df$y.south <- buff.locations.utm[,"Y"]  # UTM Northing

# Remove geometry from the main data frame
schisto_spatial_df$geom <- NULL

# ----- OPTIONAL: ----- Filtering out negative results so that we ONLY have positives to model schisto burden
# This is for when CAA is < 20
positive_only_schisto_df <- schisto_spatial_df %>%
  dplyr::filter(CAA.Estimate >= 20)
# Note: this leaves us with only 130 observations

# Can consider randomly imputing values for the 449 negative results (between 1 and 20) so that we don't lose all of our negative results in the process of analysis

# ----- OPTIONAL: ----- creating a unique location identifier and extracting one per unique location
#schisto_subset_data <- schisto_spatial_df %>%
  #mutate(location_id = paste(x.east, y.south, sep = "_")) %>%
  #group_by(location_id) %>%
  # Randomly select one buffalo per location
  #slice_sample(n = 1) %>%
  # Remove grouping
 # ungroup()

# Read the Kruger boundary - can use newer SANParks boundary - this differs only slightly from the boundary Henri has (its bounding box dimensions are xmin: 283091 ymin: 7175416 xmax: 402511 ymax: 7529638)
kruger_shape <- st_read("/Users/eberleyarborough/Documents/2023 Buffalo XS Disease Data/SANParks KNP Shapefiles/KNP_polygon/KNP_polygon.shp")

# Transform to UTM Zone 36S (EPSG:32736)
kruger_utm <- st_transform(kruger_shape, crs = 32736)

# Extract coordinates as a matrix
boundary_coords <- st_coordinates(kruger_utm)[, 1:2]

# Check the boundary coordinates are correct
cat("Boundary X range:", range(boundary_coords[,1]), "\n")
cat("Boundary Y range:", range(boundary_coords[,2]), "\n")

# Create INLA boundary segment - NO inla.mesh.map()!
bdry.utm <- inla.mesh.segment(loc = boundary_coords)

# Verify the boundary in the segment
cat("bdry.utm$loc X range:", range(bdry.utm$loc[,1]), "\n")
cat("bdry.utm$loc Y range:", range(bdry.utm$loc[,2]), "\n")

```

Simple data visualization and transformation:

Visualization code is currently commented out (as of 2/17), but this section **should** be updated if/when covariates change! For now, run this chunk for the transformations only.

Last updated 2/19 with EVI values and standardized (not logit transformed) NDVI and EVI values.

```{r data-visualization-transformation}
# ----- STEP 1: ----- Load custom visualization functions from the Highland Statistics script
source("~/Documents/Coding Help/INLA tutorials/highland_stats_INLA_course/AllRCodeInlabru/HighstatLibV14.R")

# ----- STEP 2: ----- Make a series of histograms of the covariates
# First, convert data from wide to long format
#df_long <- positive_only_schisto_df %>%
  #pivot_longer(cols = c(age.years, BCS.Average, elevation, fall.2023.NDVI, spring.2022.NDVI, fall.2023.LST.day.mean, fall.2023.LST.night.mean, fall.23.EVI, spring.22.EVI), 
               #names_to = "variable",
               #values_to = "value")

# Create multi-panel histogram
#cov_his <- ggplot(df_long, aes(x = value)) +
            #geom_histogram(bins = 30, fill = "steelblue", color = "black") +
            #facet_wrap(~ variable, scales = "free") +  # scales = "free" allows different x/y axes
            #theme_minimal() +
            #labs(title = "Distribution of Covariates",
                 #x = "Value",
                 #y = "Frequency")
#cov_his

# Age is right-skewed and should be log-transformed
# NDVI is generally bounded between 0 and 1 (although not always), so could be logit transformed or standardized
# Elevation and LST (day/night) should be standardized using the `scale()` function.

# ----- STEP 3: ----- Look for collinearity
# Making a list of covariates to plot - start with demographics
#demo_var_to_plot <- c("age.years", "BCS.Average") # can't use sex since is a categorical variable

# Making a list of environmental covariate to plot
#env_var_to_plot <- c("fall.2023.LST.day.mean", "fall.2023.LST.night.mean", "elevation", "fall.2023.NDVI", "spring.2022.NDVI", "fall.23.EVI", "spring.22.EVI")

# Using the `Mypairs` function from Highland Statistics
#age_vs_bcs <- Mypairs(positive_only_schisto_df[,demo_var_to_plot])
#env_covars_vs <- Mypairs(positive_only_schisto_df[,env_var_to_plot])

# No issues between age and BCS
# Possible issues with environmental covariates - LST day and fall NDVI have a correlation coefficient of 0.77
# Fall '23 and spring '22 NDVI have a correlation coefficient of 0.59
# LST day and LST night have a correlation coefficient of 0.53

# ----- STEP 4: ----- Transform covariates
# log transforming age using the log1p transformation to avoid taking the log of 1 (which = 0)
positive_only_schisto_df$log.age <- log1p(positive_only_schisto_df$age.years)

# logit transforming spring '22 and fall '23 NDVI
library(boot)
positive_only_schisto_df$fall.23.NDVI.logit <- logit(positive_only_schisto_df$fall.2023.NDVI)
positive_only_schisto_df$spring.22.NDVI.logit <- logit(positive_only_schisto_df$spring.2022.NDVI)

# Note: The histograms of both still show a left skew (fall '23) and a bimodal distribution (spring '22)

# Standardizing environmental covariates
# the [,1] after the `scale()` function makes R return a vector instead of a matrix (R typically returns a matrix)
positive_only_schisto_df$elevation.std <- scale(positive_only_schisto_df$elevation)[,1]
positive_only_schisto_df$LST.day.std <- scale(positive_only_schisto_df$fall.2023.LST.day.mean)[,1]
positive_only_schisto_df$LST.night.std <- scale(positive_only_schisto_df$fall.2023.LST.night.mean)[,1]
positive_only_schisto_df$fall.23.NDVI.std <- scale(positive_only_schisto_df$fall.2023.NDVI)[,1]
positive_only_schisto_df$spring.22.NDVI.std <- scale(positive_only_schisto_df$spring.2022.NDVI)[,1]
positive_only_schisto_df$fall.23.EVI.std <- scale(positive_only_schisto_df$fall.23.EVI)[,1]
positive_only_schisto_df$spring.22.EVI.std <- scale(positive_only_schisto_df$spring.22.EVI)[,1]

```

### Setting INLA values - priors, mesh values, etc.

This code chunk defines all spatial modeling parameters (e.g., mesh resolution, spatial priors) before running the models.

```{r inla_lists_setup}
###
# Setting priors and mesh values
###

# Including the KNP boundary so that the mesh is 'buffered' around the edges of the park 

#### ! Use meshbuilder() to get an idea of how these affect the mesh, If you look at the code tab there, you could copy that and replace it with our boundary and location data 

# ===== COORDINATE SYSTEM SETTINGS =====
LOCUTMset = TRUE # True = Coordinates are already in UTM (no transformation needed)
toUTMset = FALSE # (Currently unused - legacy parameter)

# ===== SPATIAL PRIOR SETTINGS =====
spdePriorFlat_set = FALSE # False = use spde priors or PC (penalized complexity) priors; true = flat or non-informative priors
# False or PC priors is better for sparse data, penalizes overly complex spatial patterns

# **Range Prior**: How far does spatial correlation extend?
# Can use the mod3 INLA range output figure to help decide what a reasonable range might be, too
spdepriorrange_set = c(7500, 0.5) # Context that may inform this prior: buffalo home ranges, river systems, local habitat similarities

# **Sigma Prior**: How much does the spatial field vary?
# on the scale of the response variable (so consider log-transformed resp vars)
spdepriorsigma_set = c(3, 0.1) # On a log scale, sigma of 5 is fairly large

# ===== MESH RESOLUTION SETTINGS =====
# These control the trade-off between spatial detail and computation speed

# *Max Edge**: Triangle side length [inner region, outer buffer]
Max_edge_set <- if(isTRUE(LOCUTMset)){
  c(5000, 20000) # 5km triangles inside boundary, 20km in outer extension
  # 5km resolution captures local-scale patterns (e.g., rivers, habitat patches)
  # 20km outer edges reduce nodes in less important buffer region
} else {
  c(0.045, 0.09) # Lat/long equivalent (if coordinates weren't in UTM, which they should be)
}

# **Cutoff**: Minimum distance between mesh nodes
Cutoff_set <- if(isTRUE(LOCUTMset)){
  c(2000) # Nodes must be greater than/equal to 2km apart
  # Prevents mesh from being too dense
  # At present, with 360 observations over ~20,000km sq, 2km spacing is reasonable
} else {
  c(0.0045) #Lat/long equivalent
}

# **Offset**: How far to extend mesh beyond the boundary [inner, outer]
Offset_set <- if(isTRUE(LOCUTMset)){
  c(3000, 5000)    # 3km inner buffer, 5km outer buffer
  # Inner buffer: ensures smooth spatial field near boundary
  # Outer buffer: handles edge effects, allows extrapolation beyond the border
  # Smaller offsets = fewer nodes = faster computation
} else {
  c(0.045, 0.009) # Lat/long equivalent
}
  
```

### Henri's Helper Functions

These are a couple of functions Henri wrote to help construct the various INLA components. These must be loaded in some fashion in order to run the models later on.

\begin{itemize}
  \item \textbf{Subset Function:} a function to basically only include rows with complete values for certain variables (somewhat redundant, as subsetting options are included within the model code)
\end{itemize}

```{r subset_function}
### Data subset function
my_dt_subset_function <- function(DT = NULL, Covar = NULL, Resp = NULL){
  dt.sub <- DT[complete.cases(DT[,colnames(DT) %in% Covar]),] 
  dt.sub <- dt.sub[complete.cases(dt.sub[,colnames(dt.sub) %in% Resp]),]  
  dt.sub
}
```

\begin{itemize}
  \item \textbf{Location Format Function:} a function that transforms any coordinates from lat/long to UTM if necessary, returns them as a matrix
  \item Necessary when coordinates aren't already in UTM (if LOCUTM = FALSE) - recall that LOCUTMset = TRUE earlier, meaning that we \textit{already} set the coordinates to UTM; in the later code chunks, LOCUTM = ? is asking if we \textit{need} to transform to UTM
\end{itemize}

```{r location_function}
### Location format function
my_location_function <- function(DT = NULL, LOCUTM = FALSE){
  if(isFALSE(LOCUTM)){
    require(sp)
    
    c.utm <- as.data.frame(cbind(DT$x.east, DT$y.south))
    names(c.utm) <-  c("X","Y")
    
    coordinates(c.utm) <- c("X","Y")
    proj4string(c.utm) <- CRS("+proj=longlat +datum=WGS84")
    
    c.utm <- spTransform(c.utm, CRS("+proj=utm +zone=36 ellps=WGS84"))
    
    # Extracting the coordinates themselves
    LOC <- coordinates(c.utm)  
    
  } else {
    LOC <- cbind(DT$x.east, DT$y.south)
  }
  
  return(LOC)
}
```

\begin{itemize}
  \item \textbf{Spatial Mesh Function:} a function that constructs the spatial mesh using the subset data, GPS coordinates, and mesh resolution values that have been run through the above functions; this mesh provides the "basis functions" used to approximate the continuous spatial surface over which we're modeling (effectively the "resolution" of the spatial field)
  \item If the `max.edge` is significantly larger than the estimated spatial range (the distance at which points stop being correlated), the mesh will be too coarse to "see" a spatial pattern; conversely, a significantly smaller `max.edge` only increases computational time/effort without providing much new information to the model
  \item \textbf{Key Mesh Parameters:}
  \begin{itemize}
    \item loc: Observation locations - the mesh will include these points
    \item boundary: Study area boundary (e.g., Kruger border) - constrains mesh extent
    \item max.edge: Controls triangle size (resolution)
    \begin{itemize}
      \item First value: max edge length INSIDE the boundary (smaller = finer resolution)
      \item Second value: max edge length in OUTER extension (can be coarser)
      \item Example: c(5000, 30000) = 5km triangles inside, 30km outside
    \end{itemize}
    \item cutoff: Minimum allowed distance between mesh nodes
    \begin{itemize}
      \item Prevents nodes from being too close together
      \item Larger cutoff = fewer nodes = faster computation
      \item Example: 3000 = nodes must be at least 3km apart
    \end{itemize}
    \item offset: How far to extend the mesh beyond the boundary
    \begin{itemize}
      \item First value: inner extension (moderate buffer)
      \item Second value: outer extension (wide buffer to handle edge effects)
      \item Example: c(10000, 30000) = 10km inner buffer, 30km outer buffer
      \item Extensions help avoid edge effects in spatial modeling
    \end{itemize}
  \end{itemize}
\end{itemize}

```{r mesh_function}
####
# Mesh function
####
my_mesh_function <- function(DT = NULL,          # Data frame with observations
                             Covar = NULL,       # Covariate names (for subsetting)
                             Resp = NULL,        # Response variable name (for subsetting)
                             LOCUTM = FALSE,     # Are coordinates in UTM or do they need conversion? (T/F)
                             toUTM = NULL,       # Currently unused parameter, not sure what Henri did with this
                             BNDRY = NULL,       # Boundary object defining the study area limits (here is `kruger_shape`)
                             Max_edge = NULL,    # Maximum triangle edge length [inner, outer], set previously
                             Cutoff = NULL,      # Minimum distance between edge nodes, set previously
                             Offset = NULL,      # Extension beyond boundary [inner, outer], set previously
                             DTsubset = FALSE){  # Should the data be subset to complete cases? (optional)
  
  # ===== STEP 1: OPTIONAL DATA SUBSETTING =====
  # If DTsubset=TRUE, remove rows with missing values in covariates or response (presumably we've already done this)
  # Actual subsetting is taken care of with the `my_dt_subset_function` code
  # This ensures the mesh is built only for locations with complete data
  if(isTRUE(DTsubset)){
  # Removes NAs so mesh nodes are only placed where we have usable data
  DT <- my_dt_subset_function(DT, Covar = Covar, Resp = Resp)
  } 
  
  # ===== STEP 2: EXTRACT LOCATION COORDINATES =====
  # Get X,Y coordinates for all observations
  # These are the "knots" - actual data points where observations were made
  # The mesh will be constrained to cover these locations plus surrounding area within KNP
  LOC <- my_location_function(DT = DT, LOCUTM = LOCUTM)
  
  # Debug output to verify coordinates are in the expected range
  # For KNP UTM: X should be ~280k - 400k, Y should be ~7.1M - 7.5M
  cat("LOC in mesh function - X range:", range(LOC[,1]), "\n")
  cat("LOC in mesh function - Y range:", range(LOC[,2]), "\n")
  
  # ===== STEP 3: VERIFY BOUNDARY COORDINATES =====
  # Check that the boundary coordinates match the data coordinate system to make sure everything is projecting correctly
  # Mismatched coordinate systems will create an invalid mesh
  if(!is.null(BNDRY)) {
    cat("BNDRY$loc - X range:", range(BNDRY$loc[,1]), "\n")
    cat("BNDRY$loc - Y range:", range(BNDRY$loc[,2]), "\n")
  }
  
  # ===== STEP 4: CREATE THE MESH =====
  # This is the core function that builds the triangular mesh
  # NOTE: fmesher package replaced the old INLA mesh functions
  # fm_mesh_2d_inla is the modern equivalent of inla.mesh.2d
 Mesh = fmesher::fm_mesh_2d_inla(
                      loc = LOC,                               # Observation coordinates (n x 2 matrix)
                      boundary = fmesher::fm_as_segm(BNDRY),   # Convert boundary to fmesher format
                      max.edge = Max_edge,                     # Triangle size control [inner, outer]
                      cutoff = Cutoff,                         # Minimum node spacing
                      offset = Offset)                         # Boundary extensions [inner, outer]
  
  # ===== STEP 5: VERIFY MESH COORDINATES =====
  # Check that mesh was created in the correct coordinate system
  # Mesh coordinates should match LOC and BNDRY coordinates
  cat("Created mesh - X range:", range(Mesh$loc[,1]), "\n")
  cat("Created mesh - Y range:", range(Mesh$loc[,2]), "\n")
  
  # ===== STEP 6: RETURN THE MESH =====
  # The mesh object contains:
  # $loc, the coordinates of all mesh vertices (nodes)
  # $graphh, the triangulation structure (which nodes connect to which)
  # $n, the total number of mesh nodes
  # This mesh will be used to create the SPDE model and projection matrix (A)
  Mesh
}

### END FUNCTION ###
```

\begin{itemize}
  \item \textbf{SPDE Function:} a function that creates a Stochastic Partial Differential Equation (SPDE) model, which defines how the spatial field varies across the study area (KNP)
  \item Specifies the spatial correlation structure with a Matérn covariance (aka continuous "spatial glue," describes how the correlation between 2 points decays as the distance between them increases) can be expressed as a \textit{solution} to an spde
  \item The spde equation is solved using the mesh constructed above, creating a sparse Gaussian Markov Random Field (GMRF, a type of mathematical "engine") approximation of a more continuous (and computationally challenging) Matérn process
\end{itemize}

```{r spde_function}
####
# spde function - here is where we set SPDE priors and controls
###

my_spde_function <- function(Mesh, spdePriorFlat = TRUE,
                             spdepriorrange = NULL,
                             spdepriorsigma = NULL){
  
  # ===== OPTION 1: FLAT (VAGUE) PRIORS =====
  # Uses default INLA priors - less informative, lets data drive estimates
  # Good for when you have no prior knowledge about spatial patterns that might be in the system
  # Drawback can lead to overfitting or unrealistic spatial range estimates  
  if(isTRUE(spdePriorFlat)){
  buff.Hosts.spde = inla.spde2.matern(mesh = Mesh)
  # Creates Matérn SPDE model with default priors
  # Matérn covariance is the standard choice for spatial modeling
  # Assumes spatial correlation decays smoothly with distance
  }
  
  # ===== OPTION 2: PENALIZED COMPLEXITY (PC) PRIORS =====
  # Uses informative priors based on prior knowledge of the study system
  # PC priors penalize complexity - in stats, it's good practice to prefer simpler models unless data strongly suggests otherwise
  # Good for when you have scientific knowledge about the spatial scale of whatever processes you're trying to model
  # This bit of code says what to do if the spdePriorFlat is FALSE - use the priors we set earlier in our INLA list setup!
  else {
  buff.Hosts.spde = inla.spde2.pcmatern(
    mesh = Mesh, 
    prior.range = spdepriorrange,     # Prior for spatial range parameter (spatial correlation decay distance)
    prior.sigma = spdepriorsigma)     # Prior for field variance parameter (marginal std. dev of spatial field, aka size of variability)
  }
  
  return(buff.Hosts.spde)
}

### END FUNCTION ###
```

**Understanding the Parameters**

\begin{itemize}
  \item Spatial range (p): this defines the distance at which the spatial correlation becomes weak
  \begin{itemize}
    \item Aka, "How far apart can two locations be before they're independent?"
    \item Small ranges (e.g., 2km) mean that spatial patterns change rapidly and are very localized
    \item Large ranges (e.g., 50km) mean that spatial patterns are broad, and values may be similar even when far apart in space
    \item In this system, we might expect intermediate ranges (5-20km) because animals move locally or broadly, depending on resource availability; however, rivers/habitats affect local areas similarly
  \end{itemize}
  \item Prior range specification: given in the format c(p0, p<p), ex: c(10000, 0.5)
  \begin{itemize}
    \item This means that for a range of 10,000m or 10km (p0), there is a 50% probability that the range is less than 10km
    \item Aka "We think spatial correlation decays within 10km, but we're not certain - it could be more or less"
  \end{itemize}
  \item Sigma: this is the marginal standard deviation of the spatial random field, or how much spatial variability exists within the study area
  \begin{itemize} 
    \item A small sigma (e.g., 0.1) means that the spatial field has little variability - the spatial effect is weak
    \item A large sigma (e.g., 2.0) means that the spatial field varies a lot - the spatial effect is strong
    \item If the response variable is on a log scale, a sigma of 0.5 means that the locations can differ by 0.5 log units due to spatial effects alone
  \end{itemize}
  \item Prior sigma specification: given in the format c(sigma, p > sigma), or c(1, 0.1)
  \begin{itemize}
    \item At a sigma of 1, there is a 10% probability that sigma exceeds 1
    \item Aka, "We think the spatial variability is moderate (<1) because we only think there is a 10% chance it's very large"
  \end{itemize}
\end{itemize}
\begin{itemize}
  \item \textbf{Model Matrix Function:} a function to make the model matrix that converts all integers and factors to characters, creates a scaffolding for a matrix without an intercept, drops the first column (becomes the reference) to avoid multicollinearity - creating dummy variables for the other categories associated with that covariate - and produces a data frame ready to be used for modeling later on (used in the `inla.stack()` command), typically as a "1" in the A list of `inla.stack`
  \item This is similar to MLR, in which the model matrix `X` maps the coefficients directly to the observations since those covariates are measured at the exact same locations as the response variable
  \item This is necessary for SPDE models because you have to manually specify the covariate structure, and with SPDE you drop your base (reference) levels because you will also manually specify an explicit intercept for the model
\end{itemize}

```{r modelmatrix_function}
#
# FUNCTION: my_modelmatrix
#
# used in Stack generation below
my_modelmatrix_function <- function(DT = NULL, Covar = NULL, CatVarSub = NULL){

# Convert any factors or integers to characters (so dummy vars are created in the matrix) 
# Note: this creates ONE, BINARY covariate per each character level - e.g., SexMale, SexFemale for `Sex`, which has 2 levels
# This is because in Bayesian stats, it's best practice to include a column for each level, with EACH representing that level's deviation from the global mean (i.e., the intercept)
# E.g., `SexMale` is the deviation from the mean associated with males, and `SexFemale` is the deviation from the overall mean associated with females
DT[,Covar] %>% dplyr::mutate(across(where(is.factor), as.character),
                             across(where(is.integer), as.character)) -> DT.modmatrix
  
X0 <- model.matrix(as.formula(paste0(" ~ -1 + ", paste(Covar, collapse = " + "))),
                   data = DT.modmatrix)

# Remove columns that are literally "(Intercept)" if they exist
X <- as.data.frame(X0[, !colnames(X0) %in% "(Intercept)"])
X
}

# Testing the output to make sure no covariates are dropped
#X_test <- my_modelmatrix_function(DT = positive_only_schisto_df, Covar = covar_mod2)
#colnames(X_test)

### END ###
```

Now that we have the model matrix and our projection (A) matrix, it's time to define a function for making the INLA stack. This is how the model expects to "see" all of the components.

\begin{itemize}
  \item \textbf{Data Stack Function:} a function that stacks all of the data - the response variable, covariates, and spatial indices together; this function should be updated if any effects (random, mixed, etc.) are changed so the model knows what to pull.
  \item The `inla.stack` function also allows you to include "NA" values for locations where you want the model to predict values in your prediction stack
  \item NOTE: the `inla.stack()` function requires \textbf{non-empty A and effects lists} to construct the data structure properly. For Model 1, which contains only fixed effects (no random or spatial effects), there is a minimal placeholder index (idx = 1:n) in the effects list with an identity projection matrix - this placeholder is NOT included in the model formula and has no impact on model estimation; it just satisfies INLA's structural requirements for including that component of the stack object
\end{itemize}

```{r stack_function}
####
# Stack function
####

my_stack_function <- function(
    DT = NULL, 
    Mod.Matrix = NULL, 
    Covar = NULL, 
    Resp = NULL, 
    GroupVar = NULL, 
    spdePriorFlat = FALSE, 
    spdepriorrange = NULL, 
    spdepriorsigma = NULL, 
    LOCUTM = NULL, 
    Mesh.list = NULL, 
    BNDRY = NULL,  
    Max_edge = Max_edge_set, 
    Predict = FALSE,         # Where you update when you want the model to generate the prediction outputs
    include_section = NULL,  # Where you'd want to update if your random effect (in this case, section) changes
    include_spatial = FALSE){ # Default is set to false so spde elements don't get included in models 1 and 2
  
  # ===== STEP 1: GET LOCATION COORDINATES =====
  # Extract spatial coordinates (X, Y) from data
  # If LOCUTM=TRUE, transforms lat/long to UTM; if FALSE, uses existing coordinates
  LOC <- my_location_function(DT = DT, LOCUTM = LOCUTM)
  
  # ===== OPTIONAL: CREATE ID VARIABLES FOR RANDOM EFFECTS =====
  # See Henry's code `_hjc` for an example
  
  # ===== STEP 3: PREPARE THE COVARIATE MATRIX =====
  # Convert the model matrix "X" (which contains covariates like Sex, BCS, age, etc.) from a matrix into a list of individual named columns
  # This allows INLA to reference covariates by name (e.g., "SexFemale")
  # The model matrix was created outside this function in the `my_modelmatrix_function`
  if(!is.null(Mod.Matrix)) {
    X_list <- as.list(as.data.frame(Mod.Matrix))
    names(X_list) <- colnames(Mod.Matrix)
    } else {
      X_list <- list()
    }
  
  # ===== STEP 4: BUILD EFFECTS LIST FOR ALL MODELS =====
  # Here we put all of the fixed effects into one list
  # If a variable is in the formula, it must live in effects, not data
  fixed_effects <- c(
    list(Intercept = rep(1, nrow(DT))),
    X_list
    )
  
  # Start with fixed effects as FIRST element
  # Effects list = what lives where
  effects_list <- list(fixed_effects)
  
  # ===== STEP 6: INITIALIZE A LISTS =====
  # A list = how the things in the effects list are projected onto the observations
  A_list <- list(1)  # ONE projection for ALL fixed effects
  
  # ===== STEP 5: (OPTIONAL) ADD HERD-LEVEL RANDOM EFFECT =====
  # This code is set up to include park section as a random effect
  if(include_section) {
    NSection <- as.numeric(as.factor(DT$Section))
    effects_list <- c(effects_list, list(list(section.num = NSection)))  # SECOND element
    A_list <- c(A_list, list(1))  # SECOND identity projection for section
    }

  # ===== STEP 7: ADD SPATIAL COMPONENT (ONLY IF REQUESTED)
  # This is where spatial random effects are added via SPDE
  if(include_spatial) {
    
    # Get location coordinates again (redundant, but ensures correct coordinates)
    LOC <- my_location_function(DT = DT, LOCUTM = LOCUTM)
    Mesh = Mesh.list
  
    # ===== CREATE PROJECTION MATRIX A =====
    # buff.HostsA is the KEY matrix that projects observation locations onto the mesh (not to be confused with the A matrix earlier)
    # It's a sparse matrix: [n_observations × n_mesh_nodes]
    # Each row corresponds to one observation; columns are mesh vertices
    # Values indicate how much each mesh node contributes to that observation
    if(!is.null(GroupVar)){
    # If there's a grouping variable (e.g., year), create grouped spatial effects (we don't have this currently)
      buff.HostsA <- inla.spde.make.A(Mesh, 
                                    loc = LOC, 
                                    group = as.numeric(as.factor(DT[,GroupVar])))
      } else {
      # Standard case: just project locations onto mesh
      buff.HostsA <- inla.spde.make.A(Mesh, loc = LOC)
      }
  
     cat("buff.HostsA dimensions:", dim(buff.HostsA), "\n")
  
    # ===== CREATE SPDE MODEL =====
    # This defines the spatial correlation structure
    # Pulls in our previously specified priors for range (how far spatial  autocorrelation extends) and sigma (field variance)
     buff.Hosts.spde <- my_spde_function(Mesh,
                                        spdePriorFlat = spdePriorFlat,
                                        spdepriorrange = spdepriorrange,
                                        spdepriorsigma = spdepriorsigma)

    # ===== CREATE SPATIAL FIELD INDEX =====
    # s.index assigns an index number to each mesh node
    # This creates the "spatial.field" variable that can be used in formulas
    # Extract only the FIRST component (see below)
    s.index <- inla.spde.make.index(name = "spatial.field",
                                    n.spde = buff.Hosts.spde$n.spde)
  
    # Extract only the spatial field indices (not group or replicate components) and add to effects list as another element
    effects_list <- c(effects_list, list(spatial.field = s.index$spatial.field))
  
    # Add the spatial projection matrix `buff.HostsA` to `A_list`
    # This is the actual spatial component of the projection matrix now that is being COMBINED with the identity projection list we made earlier
    A_list <- c(A_list, list(buff.HostsA))
    }
  
 # ===== STEP 9: DEBUG OUTPUT =====
  cat("\n=== INSIDE STACK FUNCTION - PRE-STACK ===\n")
  cat("nrow(DT):", nrow(DT), "\n")
  cat("length(y):", length(DT[,Resp]), "\n")
  #cat("\nA_list length:", length(A_list), "\n")
  cat("effects_list length:", length(effects_list), "\n")
  if(length(effects_list) > 0) {
    cat("effects_list names:", paste(names(effects_list), collapse=", "), "\n")
  }
  cat("=========================================\n\n")

  # ===== STEP 10: BUILD THE INLA STACK =====
  # The INLA stack packages together 3 things:
  # The data, or the response variable (y)
  # The A list, or the list of projection matrices (how to map effects to observations)
  # The effects, or a list of all predictor variables and indices
  # INLA uses this "stack" structure to understand the model's data organization
  
  stack_data <- list(y = DT[,Resp])  # the stack data contains ONLY the response variable, `y`

  stack_model <- inla.stack(
    data = stack_data,                # Only `y`
    A = A_list,                       # All projection matrices
    effects = effects_list,           # All identity projections (if included)
    tag = "model_stack"               # Name tag for this specific stack (as opposed to "predicted stack")
    )
  
  cat("\n=== DEBUGGING STACK INTERNALS ===\n")
  cat("Checking effects_list before stack creation:\n")
  
  cat("\n=== IMMEDIATELY AFTER STACK CREATION ===\n")
  test_data <- inla.stack.data(stack_model)
  cat("Test extraction - y length:", length(test_data$y), "\n")
  cat("Test extraction - Intercept length:", length(test_data$Intercept), "\n")
  cat("========================================\n\n")

  if (!isTRUE(Predict)) {
  return(stack_model)
  }
  
  # ===== STEP 11: CREATE THE PREDICTION STACK (OPTIONAL) =====
  # If Predict=TRUE, create a second stack for making predictions
  # This has the same structure but with y=NA (unknown values to predict)
  if(isTRUE(Predict)){
  
    # Build prediction effects list conditionally (same structure as model effects)
    pred_fixed_effects <- c(
      list(Intercept = rep(1, nrow(DT))),
      as.list(as.data.frame(Mod.Matrix))
      )
    
    pred_effects_list <- list(pred_fixed_effects)
    pred_A_list <- list(1)
    
    # Add section (or other random effect) if it was included in the model
    if(include_section) {
        pred_effects_list <- c(pred_effects_list, list(section.num = NSection))
        pred_A_list <- c(pred_A_list, list(1))  # Use list(1), not just 1 to give it it's own identity projection
        }
  
   # Add spatial component if it was included in the model
   if(include_spatial) {
        pred_effects_list <- c(pred_effects_list, list(spatial.field = s.index$spatial.field))
        pred_A_list <- c(pred_A_list, list(buff.HostsA))
        }
  
    # Create prediction stack with NA response values
    stack_predict <- inla.stack(
      data = list(y = NA),             # Unknown values to predict
      A = pred_A_list,
      effects = pred_effects_list,
      tag = "stack_predict"            # Different tag to distinguish from the model stack
      )
  
  # Combine model and prediction stacks
  # INLA will fit the model using `stack_model` data
  # Then use `stack_predict` structure to generate predictions
  return(inla.stack(stack_model, stack_predict))
  } 
}  
    ### END of FUNCTION ###

```

### Defining model variables:

This next code chunk serves to define all variables and some settings for the INLA models, making them easy to modify in one code chunk later on.

At this point (1.30), we're trying to use the positive CAA values only and 2 environmental variables (elevation and fall 2023 NDVI).

```{r model_variables}
###
# Model variables
###
# ===== MODEL-SPECIFIC FIXED EFFECT SETS =====
# These are the fixed effects that will be included in the models and serve as predictor variables
# Update these if any fixed effects are changed!
covar_mod1 <- c("log.age",       # Log1p transformed age (years)
                "Sex",           # Categorical, male/female; will be converted to dummy variables SexFemale/SexMale in the mod matrix
                "BCS.Average")   # Continuous, body condition score

covar_mod2 <- c("log.age",
                #"Sex",
                "BCS.Average",
                "elevation.std",               # Continuous, scaled elevation (m)
                #"fall.23.NDVI.std",      # Continuous, logit transformed
                #"spring.22.NDVI.std",    # Continuous, logit transformed
                #"LST.day.std",  # Continuous, scaled mean daytime LST (C)
                "fall.23.EVI.std",
                "spring.22.EVI.std")
                #"soil_category") # Broad categories (4) from the 12 more specific ones

covar_mod3 <- c("log.age",             # Same as mod2 but re-written for consistency
                #"Sex",
                "BCS.Average",
                "elevation.std",               
                #"fall.23.NDVI.std",     
                #"spring.22.NDVI.std",    
                #"LST.day.std", 
                "fall.23.EVI.std",
                "spring.22.EVI.std")
                #"soil_category")
           
# ===== DATA TRANSFORMATION =====
# Create log-transformed response variable for modeling
# Note: CHANGE THIS if working with the full dataset!!
positive_only_schisto_df <- positive_only_schisto_df %>%
 mutate(log_CAA = log1p(CAA.Estimate))
  # log1p = log(x + 1) transformation
  # CAA.Estimate has a right-skewed distribution (negative binomial-like)
  # log1p transformation normalizes the distribution making it closer to Gaussian

# ===== RESPONSE VARIABLES =====
# List of outcome variables to model
resp <- c("log_CAA")    # Primary response: log-transformed CAA titer (indicator of schisto burden)
          #"T.C.Ratio")  # Secondary/placeholder for the time being 
# Note: Can run multiple models in a loop using this vector
# Each response gets its own model with same covariates and spatial structure          

# ===== MODEL FAMILIES =====
# Specify probability distribution for each response variable
resp.family <- c("Gaussian")    # For log_CAA, assumes normally distributed residuals, too
                 #"Gaussian")    # For secondary/placeholder
# Other options: "binomial" (presence/absence), "poisson" (counts), 
#                "negative.binomial" (overdispersed counts)
# Must match order and length of resp vector

# ===== VARIABLE GROUPINGS FOR PLOTTING =====
# Organize variables into categories for cleaner visualizations later - helps with plots/tables

# Environmental/landscape-level variables
environmental.vars <- c(
  "elevation.std",
  "fall.23.NDVI.std",               # Vegetation greenness just prior to capture
  #"LST.day.std",                     # Daytime land surface temp
  #"LST.night.std",                  # Nighttime land surface temp
  "spring.22.NDVI.std",               # Vegetation greenness the spring prior to capture 
  "soil_category")
  
# Individual-level demographic variables
demographic.vars <- c(
  "log.age",
  "Sex",
  "BCS.Average")

# ===== RESPONSE VARIABLE LABELS =====
# Human-readable names for responses (for plots and tables)
resp.names <- c(
  "Log Schisto CAA Titer")
  #"Schisto T/C Ratio"

# ===== MODEL SPECIFICATIONS SUMMARY =====
# Model structure for each response (log_CAA in this case):
#   log_CAA ~ Intercept + Demographics + Environmental covariates 
#            + f(section.num) + f(spatial.field) IF specified in the model
# 
# Fixed effects: demographics + env covars
# Random effects: section, for the time being
# Spatial effect: SPDE with Matérn covariance

# Create lookup table linking the variable names to the display names
resp.names.dt <- data.frame(
  resp = resp,              # Variable name in the data
  resp.names = resp.names   # Display name for plots/tables
  )
resp.names.dt$grouping <- NA     # Placeholder for categorizing responses later on

# ===== COLOR PALETTES FOR PLOTTING =====
# Pre-defined color schemes from RColorBrewer for consistent visualization
# Each palette is a sequential color scheme (light to dark)
pallete_list <- c("Blues", "BuGn", "BuPu", "GnBu", "Greens", "Greys", "Oranges",
                  "OrRd", "PuBu", "PuBuGn", "PuRd", "Purples", "RdPu", "Reds", 
                  "YlGn", "YlGnBu", "YlOrBr", "YlOrRd", "Blues", "BuGn", "BuPu", "GnBu",
                  "Greens", "Greys", "Oranges",
                  "OrRd", "PuBu", "PuBuGn",
                  "PuBuGn", "PuRd", "Purples", "RdPu", "GnBu", "Blues")

# Having multiple palettes allows different visualizations for different variables
# Ensures plots are distinguishable and publication-ready

### 
```

### Model Formulas

The general purpose of this chunk is to define some reusable formula building blocks for different model complexities. These strings are combined with covariate names to create complete INLA formulas. The overall structure is as follows: set[model#].[suffix] where the suffix indicates formula variant for that level of model (gaussian vs non-gaussian). The only big difference between the gaussian and non-gaussian models is how the priors and hyperparameters/hyperpriors are coded, which for now (1.30) is the same, because we have no random effects at this point in time.

The complete formula is built as:

set#.left (formula header) + fixed\_# (fixed effects for that model set) +/- random effect, either .a for gaussian or .b for non-gaussian models Example: "set1.left " + "fixed_1" Results in: y \~ -1 + Intercept + age + sex + BCS

The "-1 + Intercept" removes R's automatic intercept, and "+ Intercept" adds it back explicitly as a named variable in the stack. This gives us control over intercept priors and ensures correct INLA stack structure.

At present (1.30), these model sets are provided in increasing order of complexity from 1 - 3: set 1 includes demographics ONLY, set 2 includes demographics, environmental variables, and a random effect of management section, and set 3 includes the effects from sets 1 & 2, with a spatial component. These should be updated if any changes to effects occur, as well.

```{r inla-model-formulas}
####
# Sets of formulas1-3, each with a and b
####

# ===== DEFINE UNIVERSAL MODEL COMPONENT =====
# This is the "left" side - the response and intercept specification which will be in every model 
form.left<- paste0("y ~ -1 + Intercept + ")

# ===== MODEL SET 1: DEMOGRAPHICS (no random effect, no spatial effect) =====
# Simple fixed effects model
# Option to include a random effect, ex: f(animal.numid, model='iid') = independent random intercepts per animal
# See `_hjc` code for example of how to include a random effect and what types of different priors you need for a gaussian vs non-gaussian model
# These would be named  set1.a and set1.b for the two different random effect "endings" to the model
# At this point, there is no need to define other parts of this model so there is no `set1.a/set1.b` here

# COMPLETE MODEL 1 FORMULA (Gaussian):
# form.left: y ~ -1 + Intercept +
# colnames from Xsub.filter: SexFemale + SexMale + age.years + BCS.Average

# ===== MODEL SET 2: ENVIRONMENTAL VARIABLES (no spatial) =====
# Currently adds environmental variables park management section as a random effect
# Tests whether disease varies by geographic management unit
# Assumes sections are exchangeable (no spatial structure between them)
# Uses the same form.left as model 1

# Adds in a random intercept for each park section
set2.a <- paste0(" +  f(section.num, model = 'iid')") 
# section.num gives a numeric ID for each section (Malelane, Crocodile Bridge, etc.)
# Assumes some degree of variance between sections, but no spatial correlation
# Interpretation of this effect would be "Some sections have generally higher/lower disease, independent of location"
# Can also include another random effect that may differ between gaussian/non-gaussian models - see `_hjc` code for example

# Non-Gaussian version of set a is the same
set2.b <- paste0(" +  f(section.num, model = 'iid')") 

# COMPLETE MODEL 2 FORMULA (Gaussian):
# form.left: y ~ -1 + Intercept + 
# colnames of Xsub.filter: SexFemale + SexMale + age.years + BCS.Average + ENV VARS + 
# set2.a:   f(section.num, model='iid')

# ===== MODEL SET 3: SPATIAL EFFECTS =====
# Adds continuous spatial random field using SPDE
# Captures smooth spatial variation across the landscape
# Most complex  set of models
# Same left side of the model as the other 2

# Spatial random field using SPDE approximation
set3.a <- paste0(" +  f(spatial.field, model = spde)") 
# spatial.field is an object containing indices linking observations to mesh nodes (from inla.spde.make.index)
# model = spde tells the model to use the SPDE object created by my_spde_function
# This contains the Matérn covariance structure, as well as the range and variance controlled by PC priors
# Hopefully, this captures smooth spatial patterns not explained by covariates
# Examples: habitat quality gradients, water availability, unmeasured environmental factors
# Again, see `_hjc` code for examples of keeping OTHER random effects with gaussian vs non-gaussian priors

# The non-gaussian version (again, only looks different with other random effects and their priors)
set3.b <- paste0(" +  f(spatial.field, model = spde)") 

# COMPLETE MODEL 3 FORMULA (Gaussian):
# form.left: y ~ -1 + Intercept + SexFemale + SexMale + age.years + BCS.Average + elevation + ENV VARS +
# set3.a:   f(spatial.field, model=spde)

### DONE - no need to include temporal/spatiotemporal effects here, but see Henri's code for more complex models
```

## Running the limited model(s)

### Model 1: Demographic Model

This is a fixed-effects model to understand the effects of host demographics on Schisto burden, given by:

y \~ -1 + Intercept + SexFemale + SexMale + age.years + BCS.Average, where y = log1p(CAA)

```{r model-1, cache=TRUE, warning=FALSE}
# ===== INITIALIZE STORAGE LISTS =====
# Store results from each response variable separately
mod1.list <- list()         # Full INLA model objects
mod1.stack.list <- list()   # Data stacks
mod1.mesh.list <- list()    # Meshes (created but not used in Model 1)
mod1.spde.list <- list()    # SPDE objects (created but not used in Model 1)

# ===== MODEL LOOP =====
# Run one model for each response variable
for(x in 1:length(resp)){
  
  # ----- Progress indicator ----- (will show in console, not in knitted doc) - tells you which model is running
  cat("Processing model 1:", resp[x], "-", x, "of", length(resp), "\n")
  
  # ----- Step 1: Prepare Data -----
  # Remove rows with missing values in response or covariates
  # Ensures complete cases for each specific response variable
  dt.sub <- my_dt_subset_function(DT = positive_only_schisto_df, 
                                   Covar = covar_mod1, # Filtering based on demographic covariates only
                                   Resp = resp[x])
  dt.sub <- as.data.frame(dt.sub)
  
  cat("✓ Data subset complete, n =", nrow(dt.sub), "\n")
  
  # ----- Step 2: Create Mesh -----
  # Note: Mesh is created but NOT used in Model 1 (no spatial component)
  # Kept here for consistency with Model 3 code structure
  mesh <- my_mesh_function(DT = dt.sub,
                           Covar = covar_mod1,
                           Resp = resp[x],
                           LOCUTM = LOCUTMset,
                           BNDRY = kruger_shape,
                           Max_edge = Max_edge_set,
                           Cutoff = Cutoff_set,
                           Offset = Offset_set,
                           DTsubset = FALSE) # Data already subset above
  mod1.mesh.list[[resp[x]]] <- mesh
  cat("✓ Mesh construction complete\n")
  
  # ----- Step 3: Create Model Matrix -----
  # Converts categorical variables (like Sex) into dummy variables
  # Result: numeric matrix where each column is a predictor
  Xsub <- my_modelmatrix_function(DT = dt.sub, 
                                  Covar = covar_mod1,
                                  CatVarSub = CatVarSub)
  cat("✓ Model matrix created, dim =", nrow(Xsub), "x", ncol(Xsub), "\n")
  
  # ----- Step 4: Create INLA Stack -----
  # Packages data in INLA's required format
  # include_spatial = FALSE means no spatial random effects in Model 1
  stack <- my_stack_function(Resp = resp[x],
                             Covar = covar_mod1,
                             DT = dt.sub, 
                             Mod.Matrix = Xsub,
                             GroupVar = NULL,
                             include_section = FALSE,   # No section random effect
                             include_spatial = FALSE,   # No spatial component unless model set 3
                             spdePriorFlat = spdePriorFlat_set,  # we set this to FALSE earlier to use our PC priors
                             spdepriorrange = spdepriorrange_set,
                             spdepriorsigma = spdepriorsigma_set,
                             LOCUTM = NULL,
                             BNDRY = kruger_shape, 
                             Max_edge = Max_edge_set,
                             Mesh.list = mesh) 
  
  mod1.stack.list[[resp[x]]] <- stack
  cat("✓ Stack created\n")
  
  # ----- Step 5: Set Priors for Residual Precision -----
  # PC prior for residual variance (how much unexplained variability remains)
  sdres <- sd(dt.sub[,resp[x]], na.rm = T)
  pcprior <- list(prec = list(prior="pc.prec", param = c(3*sdres,0.01))) 
  # this is a penalized complexity prior - helps to prevent overfitting
  # says we think the residual SD is very close to the observed SD, with only a 1% chance of being greater than 3x the observed SD
  
  cat("✓ Priors set\n")
  
  # ----- (Optional) Step 6: Filter Covariates (if needed) -----
  # We should only be using demographic covariates anyway, so is a placeholder for now
  Xsub.filter <- Xsub
  
  # ----- Step 7: Create SPDE -----
  # Note: SPDE is created but NOT used in Model 1 (no spatial component)
  # Kept for code consistency with Model 3
  SPDE <- my_spde_function(Mesh = mesh,
                           spdePriorFlat = spdePriorFlat_set,
                           spdepriorrange = spdepriorrange_set,
                           spdepriorsigma = spdepriorsigma_set)
  mod1.spde.list[[resp[x]]] <- SPDE
  cat("✓ SPDE created\n")
  
  # ----- Step 8: Build Model Formula -----
  # Create formula string based on family and covariates
  if(resp.family[x] == "gaussian"){
    modform <- as.formula(paste0(form.left, paste0(colnames(Xsub.filter), collapse = " + "))
)}
  # Results in something like:
    # y ~ -1 + Intercept + SexFemale + SexMale + age.years + BCS.Average
    # see `_hjc` code for an example with a gaussian random effect
  else {
    modform <- as.formula(paste0(form.left, paste0(colnames(Xsub.filter), collapse = " + "))
)  # what to use if non-gaussian family - right now is the same
  }
  
  cat("✓ Formula:", as.character(modform)[2], "\n")
  
  # ----- Pre-INLA Debug Code -----
  # Checking dimensions of response and covariates match
  cat("\n=== PRE-INLA DEBUG ===\n")
  cat("Formula:\n")
  print(modform)
  cat("\nStack data structure:\n")
  stack_data <- inla.stack.data(stack)
  print(names(stack_data))
  cat("y length:", length(stack_data$y), "\n")
  cat("y has NAs:", sum(is.na(stack_data$y)), "\n")

  # Check each component in stack data
  for(name in names(stack_data)) {
    if(name != "spde") {
      cat(name, "- length/dim:", 
         if(is.matrix(stack_data[[name]])) dim(stack_data[[name]]) else length(stack_data[[name]]), 
        "\n")
   }
  }
  cat("=====================\n\n")
  
  # ----- Step 9: Fit INLA Model -----
  # Run the actual Bayesian inference
  error_next = FALSE
  set.seed(20084) # For reproducibility
  
  tryCatch(
    mod.run <- inla(modform,
                    family = resp.family[x],                             # Gaussian for log_CAA
                    data = inla.stack.data(stack),                       # Data from stack, NO SPDE
                    control.compute = list(dic = TRUE, waic = TRUE),     # Compute model fit metrics
                    control.predictor = list(A = inla.stack.A(stack))),  # Projection matrix
    error = function(e) {
      error_next <<- TRUE
      cat("ERROR on:", resp[x], "\n")
      cat("Error message:", e$message, "\n")
    })
  
  # Skip to next response if this one failed
  if(error_next) {next}
  
  cat("✓ INLA model complete\n")
  
  # ----- Step 10: Store Results -----
  mod1.list[[resp[x]]] <- mod.run
}
## ===== END MODEL 1 LOOP =====


# ===== EXTRACT MODEL FIT STATISTICS =====

# ----- Extract DIC (Deviance Information Criterion) -----
# DIC: model fit metric (lower = better fit)
# Penalizes model complexity (similar to AIC)
mod1.dic.list <- list()
for(i in 1:length(resp)){
  mod1.dic.list[[i]] <- mod1.list[[i]]$dic$dic
}

# Handle cases where DIC wasn't computed properly
# Recalculate from local DIC values if needed
if(length(which(is.na(mod1.dic.list))) > 0){
  for(xna in 1:length(which(is.na(mod1.dic.list)))){
    sub <- (which(is.na(mod1.dic.list)))[1]
    mod1.dic.list[[sub]] <- sum(mod1.list[[sub]]$dic$local.dic, na.rm = TRUE)
    cat("Fixed DIC for var:", resp[sub], "\n")
  }
}

# Convert list to matrix for easy viewing
mod1.dic.list <- do.call(rbind, mod1.dic.list)

# Optional: save DIC values
#write.csv(mod1.dic.list, file.path(file_path, "inla.model","dic","mod1_dic.csv"))

# ----- Extract Model Summaries -----
# Pull out coefficients, credible intervals, etc. from each model
# myInla_extract() is a custom function (defined elsewhere) that formats INLA output - is currently edited to remove info from animal.ID
mod1.list.extract <-
  lapply(1:length(resp), function(x){
    myInla_extract(mod1.list[[x]], ModelName = resp[x])
  })

# Display summaries
print(mod1.list.extract)

# Save extracted summaries for later use
save(mod1.list.extract, file = file.path(file_path, "inla.model","mod1.list.extract.rds"))
# To reload later: load(file = file.path(file_path, "inla.model","mod1.list.extract.rds"))

rm(mod1.stack.list, mod1.mesh.list, mod1.spde.list); gc()

# ----- Model Selection -----
# Compare different covariate combinations (if desired)
# Example using custom function:
# HostModelSel <- INLAModelSel(resp[1], covar, "animal.ID", "iid", "Gaussian", Xsub.filter)

## ===== END MODEL 1 POST-PROCESSING =====r))
```

Now we'll plot some of the model ouputs. These could be included in the model loop code (and might be later), but for now, these are a separate code chunk.

```{r model-1-plotting}
### MODEL OUTPUT FIGURES

# Extract fixed effects from the model
fixed_effects_mod1 <- mod1.list$log_CAA$summary.fixed

# Add variable names as a column
fixed_effects_mod1$variable <- rownames(fixed_effects_mod1)

# Remove intercept if you want to focus on covariates only (optional)
#fixed_effects <- fixed_effects %>%
  #filter(variable != "Intercept")

# Create pretty labels for variables
variable_labels <- data.frame(
  variable = c("Intercept", "log.age", "SexFemale", "SexMale", "BCS.Average"),
  label = c("Intercept", "Age (log years)", "Sex: Female", "Sex: Male", "Avg. Body Condition Score")
  )

# Join labels to fixed effects
fixed_effects_mod1 <- fixed_effects_mod1 %>%
  left_join(variable_labels, by = "variable") %>%
  mutate(display_label = ifelse(is.na(label), variable, label))

# Create caterpillar plot
caterpillar_plot_mod1 <- ggplot(fixed_effects_mod1, aes(x = mean, y = reorder(display_label, mean))) +
  # Point estimate
  geom_point(size = 3) +
  # 95% Credible interval
  geom_errorbarh(aes(xmin = `0.025quant`, xmax = `0.975quant`), height = 0.2) +
  # Reference line at zero
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  # Labels
  labs(
    title = "Model 1: Host Demographic Effects on Schisto Burden",
    x = "95% Credible Interval",
    y = NULL
  ) +
  theme_classic() +
  theme(
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold")
  )

print(caterpillar_plot_mod1)

# Save the plot
ggsave(file.path(file_path, "figures", 
                 paste0("caterpillar_mod1_", "demographics_only", ".png")),
       caterpillar_plot_mod1, width = 8, height = 6, dpi = 300)

```

### Model 2: Demographics + Environment

This model is a mixed-effects model, including host demographics and environmental variables as fixed effects, and management section as a random effect. The model equation is given by:

y \~ -1 + Intercept + SexFemale + SexMale + age.years + BCS.Average + elevation + fall.2023.NDVI + fall.2023.mean.LST.day + soil_category + f(section.num, model='iid')

```{r model-2, cache=TRUE, warning=FALSE}
# ===== INITIALIZE STORAGE LISTS =====
mod2.list <- list()        # Full INLA model objects
mod2.stack.list <- list()  # Data stacks
mod2.mesh.list <- list()   # Meshes (created but not used without spatial)
mod2.spde.list <- list()   # SPDE objects (created but not used without spatial)

# ===== MODEL LOOP =====
for(x in 1:length(resp)){
  
  # ----- Progress Tracking -----
  cat("Processing model 2:", resp[x], "-", x, "of", length(resp), "\n")
  
  # ----- Step 1: Prepare Data -----
  dt.sub <- my_dt_subset_function(DT = positive_only_schisto_df, # currently running with only schisto positive animals
                                   Covar = covar_mod2, # host demographics and environmental covariates
                                   Resp = resp[x])
  dt.sub <- as.data.frame(dt.sub)
  
  cat("✓ Data subset complete, n =", nrow(dt.sub), "\n")
  
  # ----- Optional: Scale Response Variable -----
  # Commented out - only needed if response has very wide range
  # Z-score transformation: centers at 0, scales to SD = 1
  # Helps with numerical stability but makes interpretation harder
  # dt.sub[,resp[x]] <- scale(dt.sub[,resp[x]], scale = TRUE)
  
  # ----- Step 2: Create Mesh -----
  # Note: Mesh created for code consistency but not used (no spatial component)
  mesh <- my_mesh_function(DT = dt.sub,
                           Covar = covar_mod2,
                           Resp = resp[x],
                           LOCUTM = LOCUTMset,
                           BNDRY = kruger_shape,
                           Max_edge = Max_edge_set,
                           Cutoff = Cutoff_set,
                           DTsubset = FALSE)          # Already subset above
  mod2.mesh.list[[resp[x]]] <- mesh
  cat("✓ Mesh construction complete\n")
  
  # ----- Step 3: Create Model Matrix -----
  Xsub <- my_modelmatrix_function(DT = dt.sub, Covar = covar_mod2) #, CatVarSub = CatVarSub) if including any grouping factor
  # Note: CatVarSub parameter not needed for current covariates - may consider capture group
  
  # ----- Step 4: Create INLA Stack -----
  # KEY DIFFERENCE FROM MODEL 1: include_section = TRUE
  # This adds park section as a random effect
  stack <- my_stack_function(Resp = resp[x],
                             Covar = covar_mod2,
                             DT = dt.sub, 
                             Mod.Matrix = Xsub,
                             GroupVar = NULL,
                             include_section = FALSE,                # Adding section as a random effect
                             include_spatial = FALSE,               # Still no spatial component
                             spdePriorFlat = spdePriorFlat_set,
                             spdepriorrange = spdepriorrange_set,
                             spdepriorsigma = spdepriorsigma_set,
                             LOCUTM = LOCUTMset,
                             BNDRY = kruger_shape, 
                             #CatVarSub =  include the repeated sampling group here if needed - might include capture group here
                             Max_edge = Max_edge_set,
                             Mesh.list = mesh)
 
   cat("✓ Stack created\n")
   mod2.stack.list[[resp[x]]] <- stack
  
  # ----- Step 5: Set Priors -----
  sdres <- sd(dt.sub[,resp[x]], na.rm = T)
  pcprior <- list(prec = list(prior="pc.prec", param = c(3*sdres,0.01)))
  cat("✓ Priors set\n")
  
   # ----- Step 6: Filter Covariates (Optional) -----
  # Placeholder for removing specific variables if needed
  # Example: might want to drop "section" as a fixed effect if it was previously included in the model since it's now random
  # Currently keeps all variables
  #if(resp[x] == "condition"){
  #  Xsub %>% dplyr::select((!contains(c("herd","condition")))) -> Xsub.filter
  #} else {
  #  Xsub %>% dplyr::select(!contains(c("herd"))) -> Xsub.filter
  #}
  Xsub.filter <- Xsub
  
  # ----- Step 7: Create SPDE -----
  # Created for consistency but not used without spatial component
  SPDE <- my_spde_function(Mesh = mesh,
                           spdePriorFlat = spdePriorFlat_set,
                           spdepriorrange = spdepriorrange_set,
                           spdepriorsigma = spdepriorsigma_set)
  mod2.spde.list[[resp[x]]] <- SPDE
  cat("✓ SPDE created\n")
  
  # ----- Step 8: Build Model Formula -----
  # Formula now includes f(section.num, model='iid')
  # form.left, set2.a, set2.b defined in the model formula code chunk
  if(resp.family[x] == "gaussian"){
    modform <- as.formula(paste0(form.left, paste0(colnames(Xsub.filter), collapse = " + ")))  # currently have the random effect taken out
  } 
  # Results in something like:
  # y ~ -1 + Intercept + SexFemale + SexMale + age.years + BCS.Average + elevation +
  #     f(section.num, model='iid') - Section random effect
  else {
  modform <- as.formula(paste0(form.left, paste0(colnames(Xsub.filter), collapse = " + ")))  # random effect taken out
  }
  cat("✓ Formula:", as.character(modform)[2], "\n")
  
  # ===== PRE-INLA DEBUG CODE =====
  cat("\n=== PRE-INLA DEBUG ===\n")
  cat("Formula:\n")
  print(modform)
  
  stack_data <- inla.stack.data(stack)
  cat("\nStack data structure:\n")
  cat("Names in stack data:", paste(names(stack_data), collapse=", "), "\n")
  
  cat("\nChecking each variable in formula:\n")
  cat("'y' in stack_data:", "y" %in% names(stack_data), "\n")
  cat("'Intercept' in stack_data:", "Intercept" %in% names(stack_data), "\n")
  cat("'section.num' in stack_data:", "section.num" %in% names(stack_data), "\n")
  
  for(covar in colnames(Xsub.filter)) {
    cat("'", covar, "' in stack_data: ", covar %in% names(stack_data), "\n", sep="")
  }
  cat("=====================\n\n")
  
  # ----- Step 9: Fit INLA Model -----
  error_next = FALSE #For tryCatch
  set.seed(20084)
  
  tryCatch({
    mod.run <-
      inla(modform,
           family = resp.family[x],
           data = inla.stack.data(stack),
           control.compute = list(dic = TRUE, waic = TRUE),
           control.predictor = list(A = inla.stack.A(stack)))
   }
  , error = function(e) {
    error_next <<- TRUE
    cat("========= ERROR =========\n")
    cat("Error on:", resp[x], "\n")
    cat("Error message:", e$message, "\n")
    cat("========================\n")
  })
  
  if(error_next) {next}
  
  cat("✓ INLA model 2 complete\n")
  
  # ----- Step 10: Store Results -----
  mod2.list[[resp[x]]] <- mod.run
} 
### END Model 2 LOOP ###

# ===== POST-PROCESSING =====

print("Completed model loop, now saving objects")

# ----- Optional: Save Large Objects -----
# Uncomment to save for later reloading
#save(mod2.list, file = file.path(file_path, "inla.model", "mod2_list.R"))

# Uncomment to  load later on: load(file = file.path(file_path, "inla.model","mod2_list.R"))
#save(mod2.stack.list, file = file.path(file_path, "inla.model", "mod2_stack_list.R"))
#save(mod2.mesh.list , file = file.path(file_path, "inla.model", "mod2_mesh_list.R"))
#save(mod2.spde.list , file = file.path(file_path, "inla.model", "mod2_spde_list.R"))

# ----- Clean Up Memory -----
# Remove large objects no longer needed (keep mod2.list for analysis)
rm(mod2.stack.list, mod2.mesh.list, mod2.spde.list); gc()

cat("Completed saving objects now extracting model results...\n")

# ===== EXTRACT MODEL FIT STATISTICS =====

# ----- Extract DIC -----
mod2.dic.list <- list()
for(i in 1:length(resp)){
  mod2.dic.list[[i]] <- mod2.list[[i]]$dic$dic
}

# Handle missing DIC values
if(length(which(is.na(mod2.dic.list))) > 0){
  for(xna in 1:length(which(is.na(mod2.dic.list)))){
    sub <- (which(is.na(mod2.dic.list)))[1]
    mod2.dic.list[[sub]] <- sum(mod2.list[[sub]]$dic$local.dic, na.rm = T)
    print(paste0("Fixed DIC for var: ", resp[sub]))
  }
}

mod2.dic.list <- do.call(rbind, mod2.dic.list)
# Optional: Save DIC
write.csv(mod2.dic.list, file.path(file_path, "inla.model", "dic", "mod2_dic.csv"))

# ----- Extract Model Summaries -----
mod2.list.extract <-
  lapply(1:length(resp), function(x){
    myInla_extract(mod2.list[[x]], ModelName = resp[x])
  })

# ----- Display Regression Table -----
# Creates publication-ready table in console
# Shows coefficients, standard errors, credible intervals
texreg::screenreg(mod2.list.extract)

# Save extracted summaries
#save(mod2.list.extract, file = file.path(file_path, "inla.model","mod2.list.extract.rds"))
# To reload: load(file = file.path(file_path, "inla.model","mod2.list.extract.rds"))

# ===== VISUALIZATION: =====
# See the model 2 code for Henri's code making nice tables for his dissertation!

# ----- Caterpillar Plot for Section Random Effects -----
# Shows estimated effect for each park section with uncertainty
# Helps identify which sections have higher/lower disease burden

# Extract section random effects from the last model run
#section_effects <- mod.run$summary.random$section.num

# Create the caterpillar plot 
#section_effect_plot <- ggplot(section_effects, aes(x = mean, y = reorder(ID, mean))) +
  #geom_point(size = 3) +  # Point estimate
  #geom_errorbarh(aes(xmin = `0.025quant`, xmax = `0.975quant`), 
                # height = 0.2) +  # 95% credible intervals
  #geom_vline(xintercept = 0, linetype = "dashed", color = "red") +  # Reference line
  #labs(
    #x = "Section Effect (log scale)",
    #y = "KNP Management Section",
    #title = "Random Effects by Park Section",
    #subtitle = "95% Credible Intervals"
  #) +
  #theme_minimal()
#section_effect_plot

# INTERPRETATION:
# - Points to the right of 0 = higher disease in that section
# - Points to the left of 0 = lower disease in that section  
# - Wide intervals = high uncertainty (few observations in that section)
# - Narrow intervals = more confident estimate
# - If all intervals overlap 0 = no strong section differences
# Note: Plot shows minimal differences between sections for this dataset

# ----- Caterpillar Plot for  Fixed Effects -----
# Extract fixed effects from the model
fixed_effects_mod2 <- mod2.list$log_CAA$summary.fixed
fixed_effects_mod2$variable <- rownames(fixed_effects_mod2)

variable_labels_mod2 <- data.frame(
  variable = c("Intercept", "log.age", "BCS.Average", "elevation.std", "fall.23.EVI.std", "spring.22.EVI.std"),
  label = c("Intercept", "Age [log(years)]", "Avg. BCS", "Elevation (m)", "Fall 2023 EVI", "Spring 2022 EVI")
  )

# Join labels to fixed effects
fixed_effects_mod2 <- fixed_effects_mod2 %>%
  left_join(variable_labels_mod2, by = "variable") %>%
  mutate(display_label = ifelse(is.na(label), variable, label)) %>%
  mutate(display_label = factor(
    display_label,
    levels = c(
      "Intercept", "Age [log(years)]", "Avg. BCS", "Elevation (m)", "Fall 2023 EVI", "Spring 2022 EVI")
  ))

# Create caterpillar plot
caterpillar_plot_mod2 <- ggplot(fixed_effects_mod2, aes(x = mean, y = display_label)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = `0.025quant`, xmax = `0.975quant`), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Model 2: Demographic & Environmental Effects on Schisto Burden",
    x = "95% Credible Interval",
    y = "Fixed Effects"
  ) +
  theme_classic() +
  theme(
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold")
  )

print(caterpillar_plot_mod2)

### NOTE: The alluvial soil type ONLY HAS 1 OBSERVATION, so INLA is dropping it internally since it can't use it in the regression analysis.
# Can consider including soil type as a random effect in this model

# Save the plot
#ggsave(file.path(file_path, "figures", 
                 #paste0("caterpillar_mod2_", "demographics_and_env", ".png")),
       #caterpillar_plot_mod2, width = 8, height = 6, dpi = 300)

# ----- Computing Posterior Probabilities ----- #
# Taking each covariate and computing the posterior probability that the coefficient is NOT 0
post_prob_mod2 <- lapply(mod2.list, function(m) {
  sapply(names(m$marginals.fixed), function(nm) {
    1 - inla.pmarginal(0, m$marginals.fixed[[nm]])
    })
  })
print(post_prob_mod2)

# NOTE: change the filename for different model runs if you change covariates
#write.csv(post_prob_mod2, file.path(file_path, "inla.model", "post_prob", "mod2_evi_nosex_noLST.csv"))


# ===== MODEL COMPARISON NOTES =====
# Compare DIC values:
#   - Model 1 DIC: baseline - host demographics only, no random effects
#   - Model 2 DIC: adding in environmental covariates and park section (random effect)
#   - If Model 2 DIC is much lower (>5 points): section matters
#   - If similar: section doesn't add explanatory power
#
# Next step: Compare Model 2 vs Model 3 (spatial)
#   - If spatial model is much better: continuous spatial patterns exist
#   - If section model is better: discrete geographic differences
#   - Both could be important!

## ===== END MODEL 2 =====
```

**As of 2/17:**
Including park section as a random effect only improves DIC and WAIC by ~0.5, so it's not helpful and can be removed from the model.
Unfortunately, at this time including the environmental covariates *increases* our DIC by 1 compared to model 1.
DIC for the model including LST day but **not** NDVI is 523.1235; DIC for the model including NDVI but **not** LST day is 521.3724.


### Model 3: Full Spatial Model With SPDE

This model is a mixed-effects model and includes host demographics and environmental variables as fixed effects. This model also contains a spatial component and serves to fit models with continuous spatial random effects using INLA-SPDE. This is the most complex model of our model sets, capturing spatial autocorrelation patterns across KNP. This allows disease burden to vary smoothly across the park. The full equation is given by:

y \~ -1 + Intercept + SexFemale + SexMale + age.years + BCS.Average + elevation + fall.2023.NDVI + f(spatial.field, model=spde)

Here is a code chunk aiming to use INLA's `meshbuilder()` function to see how changes in priors, edges, and cutoff values affect the mesh, and what might fit best with this data. This function opens a `Shiny()` app, creating a new window in which you can work interactively to change the mesh, without writing any code. After you're happy with the mesh, you can copy the code over into the body of the .Rmd file.

```{r mesh-builder}
library(shiny)

#extracting buffalo locations
mesh.loc <- my_location_function(DT = schisto_spatial_df, LOCUTM = TRUE)

# Launch meshbuilder  - it opens the new window that lets you select your spatial elements individually
meshbuilder()
```

Before running the spatial model, it's good to check that the mesh works. This is an example from Henri of how the 'my_mesh_function' constructs the mesh, and how he saves it to the wd. Notice that as the triangles approach the KNP boundary, they get smaller.

```{r example_mesh}
#Mesh list
mod.mesh.list <-
#lapply(1:length(resp), function(x){ - uncomment this line to loop through all response variables listed in `resp` (in this case, two)
lapply(1:1, function(x){
my_mesh_function(DT = schisto_spatial_df,
                 Covar = covar_mod3,
                 Resp = resp[[x]],
                 LOCUTM = LOCUTMset,
                 toUTM = toUTMset,
                 BNDRY = kruger_shape, # change to Henri's boundary if issues arise
                 Max_edge = Max_edge_set,
                 Cutoff = Cutoff_set,
                 Offset = Offset_set,
                 DTsubset = TRUE)
})


plot(mod.mesh.list[[1]])
meshdt <- mod.mesh.list #[[1]]
mod.mesh.list[[1]]$n

#save(mod.mesh.list, file = file.path(file_path, "inla.model","mod_mesh_list.R"))
#load(file = file.path(file_path, "inla.model","mod_mesh_list.R"))
rm(mod.mesh.list, meshdt)
gc()

### END ###
```

Now that we've checked that the mesh works, it's time to run the spatial model where the mesh will actually be fed into the code.

```{r mod3, eval=FALSE, include=FALSE}
# ===== INITIALIZE STORAGE LISTS =====
mod3.list <- list()        # Full INLA model objects
mod3.stack.list <- list()  # Data stacks
mod3.mesh.list <- list()   # Meshes (NOW ACTUALLY USED!)
mod3.spde.list <- list()   # SPDE objects (NOW ACTUALLY USED!)
mod3.error.list <- list()  # Track any errors during model fitting

# ===== MODEL LOOP =====
for(x in 1:length(resp)){
  
  # ----- Progress Tracking -----
  cat("Processing model 3:", resp[x], "-", x, "of", length(resp), "\n")
  
  # ----- Step 1: Prepare Data -----
  dt.sub <- my_dt_subset_function(DT = positive_only_schisto_df, 
                                   Covar = covar_mod3, 
                                   Resp = resp[x])
  dt.sub <- as.data.frame(dt.sub)
  cat("✓ Data subset complete, n =", nrow(dt.sub), "\n")
  
  # ----- Optional: Scale Response Variable -----
  # Commented out - typically not needed for log-transformed data
  # dt.sub[,resp[x]] <- scale(dt.sub[,resp[x]], scale = TRUE)T)
  
  # ----- Step 2: Create Mesh -----
  # THIS IS CRITICAL - mesh defines the spatial structure!
  mesh <- my_mesh_function(DT = dt.sub,
                           Covar = covar_mod3,
                           Resp = resp[[x]],
                           LOCUTM = LOCUTMset,
                           toUTM = toUTMset,
                           BNDRY = kruger_shape,
                           Max_edge = Max_edge_set,
                           Cutoff = Cutoff_set,
                           Offset = Offset_set,
                           DTsubset = TRUE) # Creates mesh ONLY for locations where we have data
  mod3.mesh.list[[resp[x]]] <- mesh
  cat("✓ Mesh construction complete\n")

  # ----- Step 3: Create Model Matrix -----
  Xsub <- my_modelmatrix_function(DT = dt.sub, Covar = covar_mod3)#, CatVarSub = CatVarSub)
  Xsub <- as.matrix(Xsub)
 
  # Debug: Verify dimensions
  cat("Checking Xsub dimensions:\n")
  print(dim(Xsub))
  print(class(Xsub))
  
  # ----- Step 4: Create INLA Stack -----
  # KEY DIFFERENCE FROM MODELS 1 & 2: include_spatial = TRUE
  # This adds the continuous spatial random field
  stack <- my_stack_function(Resp = resp[x],
                             Covar = covar_mod3,
                             DT = dt.sub, 
                             Mod.Matrix = Xsub,
                             GroupVar = NULL,
                             include_section = FALSE, # including section as a random effect may conflict with the spatial aspect
                             include_spatial = TRUE, # ADD spatial random effect here!
                             spdePriorFlat = spdePriorFlat_set,
                             spdepriorrange = spdepriorrange_set,
                             spdepriorsigma = spdepriorsigma_set,
                             LOCUTM = LOCUTMset,
                             BNDRY = kruger_shape, 
                             #CatVarSub = "capture.year2009",
                             Max_edge = Max_edge_set,
                             Mesh.list = mesh)
  cat("✓ Stack created\n")
  mod3.stack.list[[resp[x]]] <- stack
  
  # ----- Step 5: Set Priors -----
  sdres <- sd(dt.sub[,resp[x]], na.rm = T)
  pcprior <- list(prec = list(prior="pc.prec", param = c(3*sdres,0.01)))
  cat("✓ Priors set\n")
  
  # ----- Step 6: Create SPDE -----
  # THIS MATTERS - defines the spatial covariance structure!
  SPDE <- my_spde_function(Mesh = mesh
                           , spdePriorFlat = spdePriorFlat_set
                           , spdepriorrange = spdepriorrange_set
                           , spdepriorsigma = spdepriorsigma_set)
  mod3.spde.list[[resp[x]]] <- SPDE
  cat("✓ SPDE created\n")
  
  # ----- Step 7: Filter Covariates (Optional) -----
  #if(resp[x] == "condition"){
  #  Xsub %>% dplyr::select((!contains(c("herd","condition")))) -> Xsub.filter
  #} else {
  #  Xsub %>% dplyr::select(!contains(c("herd"))) -> Xsub.filter
  #}
  Xsub.filter <- Xsub
  
  # Debug: Check covariate names
  cat("Xsub.filter column names:\n")
  print(colnames(Xsub.filter))
  cat("\nFormula being used:\n")
  
  # ----- Step 8: Build Model Formula -----
  # Formula now includes f(spatial.field, model = spde)
  # set3.a, set3.c, set3.d defined elsewhere
  if(resp.family[x] == "gaussian"){
    # Simplified formula for debugging - just intercept, covariates, and spatial
    #modform <- as.formula("y ~ -1 + Intercept + X + f(spatial.field, model = spde)")
    # Full formula would be:
    modform <- as.formula(paste0(form.left, paste0(colnames(Xsub.filter), collapse = " + "), set3.a))
    # Results in:
    # y ~ -1 + Intercept + SexFemale + SexMale + age.years + BCS.Average + elevation +
    #     f(spatial.field, model=spde) - Continuous spatial random effect!
  } else {
    modform <- as.formula(paste0(form.left, paste0(colnames(Xsub.filter), collapse = " + "), set3.b))   
  }

  cat("✓ Formula:", as.character(modform)[2], "\n")
  
  # Debug: Check stack structure
  cat("Stack data contents:\n")
  stack_data <- inla.stack.data(stack, spde = SPDE)
  print(names(stack_data))
  cat("\n")
 
  # ----- Step 9: Fit INLA Model -----
  error_next = FALSE #For tryCatch
  set.seed(20084)
  
  tryCatch({
    mod.run <-
      inla(
        modform,
        family = resp.family[x],
        data = inla.stack.data(stack, spde = SPDE),
        control.inla = list(int.strategy = "auto"),    # Let INLA choose integration strategy
        control.compute = list(dic = TRUE, waic = TRUE),
        control.predictor = list(A = inla.stack.A(stack)), 
        safe=TRUE # True - safe mode to avoid numerical issues
        )
    }, error = function(e) {
    error_next <<- TRUE
    cat("========= ERROR =========\n")
    cat("Error on:", resp[x], "\n")
    cat("Error message:", e$message, "\n")
    cat("========================\n")
    mod3.error.list[[resp[x]]] <- resp[x]
  })
  
  if(error_next) {next}
  
  cat("✓ INLA model 3 complete\n")
  
  # ----- Step 10: Store Results -----
  mod3.list[[resp[x]]] <- mod.run
  
  # Optional: Clean up memory during loop
  #rm(dt.sub, mesh, Xsub, Xsub.filter, stack, sdres, pcprior, SPDE, modform, mod.run)
  #gc()
} 

### END MODEL 3 LOOP ###

print("Completed model loop, now saving objects")

# ===== POST-PROCESSING =====

# ----- Clean Up Memory -----
# Keep mod3.list and mod3.mesh.list for analysis
# Remove large intermediate objects
#rm(mod3.stack.list, mod3.spde.list)
#gc()

# ===== EXTRACT MODEL FIT STATISTICS =====

# ----- Extract DIC -----
mod3.dic.list <- list()
for(i in 1:length(resp)){
  mod3.dic.list[[i]] <- mod3.list[[i]]$dic$dic
}

# Handle missing DIC values
if(length(which(is.na(mod3.dic.list))) > 0){
for(xna in 1:length(which(is.na(mod3.dic.list)))){
  sub <- (which(is.na(mod3.dic.list)))[1]
  mod3.dic.list[[sub]] <- sum(mod3.list[[sub]]$dic$local.dic, na.rm = T)
  print(paste0("Fixed DIC for var: ", resp[sub]))
}
}

mod3.dic.list <- do.call(rbind, mod3.dic.list)
# Optional: Save DIC
write.csv(mod3.dic.list, file.path(file_path, "inla.model", "dic", "mod3_dic.csv"))

# ----- Computing Posterior Probabilities ----- #
post_prob_mod3 <- lapply(mod3.list, function(m) {
  sapply(names(m$marginals.fixed), function(nm) {
    1 - inla.pmarginal(0, m$marginals.fixed[[nm]])
    })
  })
print(post_prob_mod3)

# NOTE: change the filename for different model runs if you change covariates
write.csv(post_prob_mod3, file.path(file_path, "inla.model", "post_prob", "mod3_evi_nosex_noLST.csv"))

# ----- Extract Model Characteristics -----

# Number of observations per model
n.samp <- 
  c(do.call(cbind,
            lapply(1:length(mod3.list), function(i){
              length(mod3.list[[i]]$residuals$deviance.residuals)
            })))

# 129

# Number of unique individuals (random effect levels)
n.ind <-
  c(do.call(cbind,
            lapply(1:length(mod3.list), function(i){
              mod3.list[[i]]$size.random[[1]]$n
            })))

# 2422

# ----- Extract Spatial Range Parameter (ρ) and precision (inverse variance) of the spatial field -----
# Rho quantifies spatial autocorrelation strength
# Here we are extracting the posterior mean and 95% CI for rho
# INLA takes precision and reports the standard deviation
# We want the sigma squared - aka the "marginal variance" of the spatial field
# σ² tells you how much variation in your response is explained by spatial structure, after accounting for all fixed effects
# σ² also is on the same scale as the response variable -  so for schisto, we will need to back-transform

# Initialize empty lists to store results
rho_list <- list()
sigma2_list <- list()

# Loop over models
for(i in seq_along(mod3.list)) {
  hyper <- mod3.list[[i]]$summary.hyperpar
  
  # Spatial range
  rho_list[[i]] <- c(
    mean = hyper["Range for spatial.field", "mean"],
    lower = hyper["Range for spatial.field", "0.025quant"],
    upper = hyper["Range for spatial.field", "0.975quant"]
  )

# Spatial variance (square of standard deviation)
  sigma2_list[[i]] <- c(
    mean = hyper["Stdev for spatial.field", "mean"]^2,
    lower = hyper["Stdev for spatial.field", "0.025quant"]^2,
    upper = hyper["Stdev for spatial.field", "0.975quant"]^2
  )
}

# Combine into data frames
rho_df <- do.call(rbind, rho_list)
sigma2_df <- do.call(rbind, sigma2_list)

# Add model IDs
rho_df <- data.frame(model = seq_along(mod3.list), rho_df)
sigma2_df <- data.frame(model = seq_along(mod3.list), sigma2_df)

# Optional: round for readability
rho_df[, -1] <- round(rho_df[, -1], 2)
sigma2_df[, -1] <- round(sigma2_df[, -1], 4)

# View results
rho_df
sigma2_df

# ===== EXTRACT MODEL SUMMARIES =====

mod3.list.extract <-
  lapply(1:length(resp), function(x){
    myInla_extract(mod3.list[[x]], ModelName = resp[x])
  })

# Save extracted summaries
#save(mod3.list.extract, file = file.path(file_path, "inla.model","mod3.list.extract.R"))
# To reload: load(file = file.path(file_path, "inla.model","mod3.list.extract.R"))

# ----- Optional: Create Publication Tables -----
# Uncomment to generate LaTeX regression tables
# texreg::screenreg(mod3.list.extract)
# texreg::texreg(mod3.list.extract,
#                booktabs = TRUE,
#                caption = "Results from spatial model (Model 3).",
#                label = "table:results_mod3",
#                file = file.path(file_path, "tables", "table_results_mod3.tex"))

# ===== SPATIAL FIELD VISUALIZATION =====

cat("now extracting field plots...")

# ----- Create Spatial Field Maps -----
# Shows estimated spatial random effects across Kruger
# Variable options: "mean" (posterior mean), "sd" (uncertainty), "0.025quant", "0.975quant"
# Rounding = TRUE bins estimates for clearer visualization)

mod3.list.plot <-
  lapply(1:length(resp), function(x){
    ggField <- my_ggField(
      mod3.list[[x]], 
      Mesh = mod3.mesh.list[[x]], 
      Str_name = "Range for ", 
      Variable = "mean",                  # Plot posterior mean of spatial field
      Res = 300,                          # Grid resolution for projection
      FitScale = "Link",                  # Plot on link scale (log scale for Gaussian)
      Colour_scale = T, 
      Palette = pallete_list[8],
      Rounding = TRUE                     # Bin values for cleaner map
    ) +
      ggtitle(resp[x]) +
      theme_classic() +
      my_plot_settings
    
    # Add KNP boundary overlay
    ggField <- ggField + 
      geom_sf(data = kruger_shape, inherit.aes = FALSE, 
              fill = NA, color = "black", linewidth = 1) 
    
    # Save underlying data
    data.table::fwrite(ggField$data, file.path(file_path, "inla.model", "dic", "dic_fieldplots",
                                                paste0("mod3_",x,"_fieldplot_dt.csv")))
    ggField 
  })

#save(mod3.list.plot, file = file.path(file_path, "inla.model","mod3_list_ggfield_plot.R"))
# To reload: load(file = file.path(file_path, "inla.model","mod3_list_ggfield_plot.R"))

# ----- Combine Field Plots into Multi-Panel Figure -----
cat("Now combining the field plots ...")

mod3_ggfield_combo <- patchwork::wrap_plots(mod3.list.plot) +
    plot_annotation(tag_levels = 'a') +     # Label panels a, b, c...
    plot_layout(guides='keep')              # Keep legends

my_ggsave_plots(plotname = mod3_ggfield_combo,
                plot_path = file.path(file_path, "figures"),
                plot_device = "png",
                height = textwidth,
                width = textheight,
                units = "mm")
#rm(mod3.list.plot)
#gc()

# ===== SPATIAL RANGE ANALYSIS =====
print("Extracting the range values ...")

# ----- Plot Spatial Correlation Decay -----
# Shows how correlation decreases with distance
# X-axis = distance, Y-axis = correlation
# Helps interpret spatial scale of disease patterns
mod3_inlarange_2.0 <- my_INLARange(
  mod3.list,
  mod3.mesh.list,
  Str_name = "Range for ",
  MaxRange = 10000,           # Plot up to 10km
  ModelNames = resp
) +
    theme_classic() +
    my_plot_settings   

my_ggsave_plots(mod3_inlarange_2.0,
                plot_path = file.path(file_path, "figures"),
                plot_device = "png",
                height = textwidth,
                width = textheight,
                units = "mm") 

# ----- Calculate Halving Range -----
# Distance at which spatial correlation drops to 50%
# Key metric for understanding spatial scale of patterns
# e.g., "Disease is correlated within 3km, independent beyond that"
mod3.list.halvingrange.dt <- my_INLARange(
  mod3.list,
  mod3.mesh.list,
  Str_name = "Range for ",
  MaxRange = 10000,
  ModelNames = resp,
  Return = "Data",       # Return data instead of plot
  Resolution = 10000)    # High resolution for precise halving distance

data.table::fwrite(mod3.list.halvingrange.dt, file = file.path(file_path, "inla.model", "mod3_list_halvingrange_dt.csv"))

# Find distance where correlation = 0.5
mod3.list.halvingrange <- mod3.list.halvingrange.dt %>%
  group_by(Model) %>%
  slice(which.min(abs(Cor.M - 0.5))) %>%       # Find closest to 50% correlation
  dplyr::mutate(Cor.M = round(Cor.M, 2)) %>%
  as.data.frame()

cat("Halving range (distance to 50% correlation):\n")
print(mod3.list.halvingrange)

#write.csv(mod3.list.halvingrange, file = file.path(file_path, "inla.model","mod3_halvingrange.csv"))

# INTERPRETATION GUIDE:
# If halving range = 5km → locations 5km apart have ~50% correlation
#                       → locations 10km apart have ~25% correlation
#                       → locations 15km apart are nearly independent
# Smaller halving range = more localized spatial patterns
# Larger halving range = broader spatial patterns

# ===== FINAL CLEANUP =====
#rm(mod3.list.extract, mod3.list, mod3.list.halvingrange, mod3.list.halvingrange.dt,
   #mod3_inlarange, mod3_ggfield, mod3.list.plot, mod3.mesh.list)
#gc()

# ----- Log Completion -----
mod.logfile <- paste0("completed mod3 run on: ", Sys.time())
data.table::fwrite(as.list(mod.logfile), append = T,
                   file = file.path(file_path, "inla.model", "logfiles", "inlarunlog_mod3.txt"))

cat("All done !!!")

## ===== END MODEL 3 =====
```

As of 2/17:
Model 3 does not have a significant improvement in DIC - is only 523.3577
The 95% CI for σ² (0.0039 – 1.447) translates to SD ≈ 0.062 – 1.203 → 

exp(0.062) ≈ 1.06, exp(1.203) ≈ 3.33 - So depending on uncertainty, the spatial effect could be very weak (~6% difference) or very strong (~3.3× difference) on the original scale.
